# Equal function rewards

All logic functions yield equal resources when performed by digital organisms.
No "abiotic" selective pressure to evolve more complex functions.

## Dependencies and setup

Load R library dependencies.

```{r}
library(tidyverse)
library(cowplot)
library(khroma)
library(rstatix)
library(knitr)
library(kableExtra)
```

Define relevant file paths.

```{r}
# Check if Rmd is being compiled using bookdown
bookdown <- exists("bookdown_build")

# Build relative path from R's working directory to what is actually the working
# directory for this experiment.
experiment_slug <- "2025-11-24-health-flat-rewards"
exp_directory <- paste(
  "experiments",
  experiment_slug,
  "analysis",
  sep = "/"
)

# Get path to the supplemental material analysis directory
supp_directory <- paste(
  "experiments",
  "manuscript-analyses",
  sep = "/"
)

# Adjust experiment directory if being knitted for bookdown build
if (bookdown) {
  exp_directory <- paste0(
    bookdown_wd_prefix,
    exp_directory
  )
  supp_directory <- paste0(
    bookdown_wd_prefix,
    supp_directory
  )
}

supp_file_prefix <- "flat-rewards"
```

Define default plot theme, plot output directory.

```{r}
# Use the cowplot theme as our default for ggplots
theme_set(theme_cowplot())
# Create directory to save plots
plot_dir <- paste(
  supp_directory,
  "plots",
  supp_file_prefix,
  sep = "/"
)
dir.create(
  plot_dir,
  showWarnings = FALSE
)

# Use vibrant theme for plot color palettes
vibrant_colors <- color("vibrant")(7)
color_pal <- c(
  vibrant_colors[["blue"]],
  vibrant_colors[["orange"]],
  vibrant_colors[["magenta"]]
)
```

## Load and preprocess data

```{r}
tasks <- c("NOT", "NAND", "OR_NOT", "AND", "OR", "AND_NOT", "NOR", "XOR", "EQU")
max_pop_size <- 100 * 100
threshold <- 0.025 * max_pop_size
```

Load summary data.

```{r}
summary_data_path <- paste(
  exp_directory,
  "data",
  "summary.csv",
  sep = "/"
)
summary_data <- read_csv(summary_data_path)
# Do any universal preprocessing
summary_data <- summary_data %>%
  mutate(
    SEED = as.factor(SEED),
    task_credit = paste0(
      "H", HOST_ONLY_FIRST_TASK_CREDIT,
      "S", SYM_ONLY_FIRST_TASK_CREDIT
    ),
    sym_condition = case_when(
      START_MOI == "0" ~ "no-syms",
      START_MOI == "1" & HEALTH_TYPE == "parasite" ~ "parasite",
      START_MOI == "1" & HEALTH_TYPE == "mutualist" ~ "mutualist",
      .default = "unknown" # precaution against unexpected parameterizations
    ),
    interaction_multiplier = as.factor(PARASITE_CYCLE_STEAL_MULTIPLIER)
  ) %>%
  mutate(
    task_credit = as.factor(task_credit),
    sym_condition = factor(
      sym_condition,
      levels = c("no-syms", "parasite", "mutualist")
    )
  ) %>%
  # Filter to just IM=4 data => that's our baseline in the paper.
  filter(interaction_multiplier == "4")
```

Load time series data.

```{r}
time_data_path <- paste(
  exp_directory,
  "data",
  "time_series.csv",
  sep = "/"
)
time_data <- read_csv(time_data_path)

# Do any universal preprocessing
time_data <- time_data %>%
  mutate(
    SEED = as.factor(SEED),
    sym_condition = case_when(
      START_MOI == "0" ~ "no-syms",
      START_MOI == "1" & HEALTH_TYPE == "parasite" ~ "parasite",
      START_MOI == "1" & HEALTH_TYPE == "mutualist" ~ "mutualist",
      .default = "unknown" # precaution against unexpected parameterizations
    ),
    interaction_multiplier = as.factor(PARASITE_CYCLE_STEAL_MULTIPLIER)
  ) %>%
  mutate(
    sym_condition = factor(
      sym_condition,
      levels = c("no-syms", "parasite", "mutualist")
    )
  ) %>%
  # Filter to just IM=4 data => that's our baseline in the paper.
  filter(interaction_multiplier == "4")
```

## Extinctions

Did host or symbiont populations go extinct in any runs?

```{r}
sym_min_pop_sizes <- time_data %>%
  group_by(SEED) %>%
  slice_min(
    OrgCounts_hosted_sym_count,
    n = 1
  ) %>%
  slice_min(
    update,
    n = 1
  )

reps_with_extinct_syms <- sym_min_pop_sizes %>%
  group_by(
    sym_condition
  ) %>%
  summarize(
    reps_with_extinct_syms = sum(OrgCounts_hosted_sym_count == 0),
    total_reps = n()
  )

host_min_pop_sizes <- time_data %>%
  group_by(SEED) %>%
  slice_min(
    OrgCounts_host_count,
    n = 1
  ) %>%
  slice_min(
    update,
    n = 1
  )

reps_with_extinct_hosts <- host_min_pop_sizes %>%
  group_by(
    sym_condition
  ) %>%
  summarize(
    reps_with_extinct_hosts = sum(OrgCounts_host_count == 0),
    total_reps = n()
  )

extinctions <- full_join(
  reps_with_extinct_syms,
  reps_with_extinct_hosts
)
extinctions <- extinctions %>%
  arrange(
    desc(reps_with_extinct_hosts), desc(reps_with_extinct_syms)
  )
extinction_table <- kable(extinctions) %>%
  kable_styling(latex_options = "striped")
save_kable(
  extinction_table,
  paste(
    plot_dir,
    "extinctions.pdf",
    sep = "/"
  )
)

extinctions
```

## Host function performance

First, quantify which functions have been performed over the course of an entire run.

```{r}
# Pivot task data long, selecting only relevant columns
host_over_time_parent_task_data <- time_data %>%
  select(
    SEED,
    sym_condition,
    interaction_multiplier,
    update,
    CurUpdate_AND_NOT_in_host_parent_org_counts,
    CurUpdate_AND_in_host_parent_org_counts,
    CurUpdate_EQU_in_host_parent_org_counts,
    CurUpdate_NAND_in_host_parent_org_counts,
    CurUpdate_NOR_in_host_parent_org_counts,
    CurUpdate_NOT_in_host_parent_org_counts,
    CurUpdate_OR_NOT_in_host_parent_org_counts,
    CurUpdate_OR_in_host_parent_org_counts,
    CurUpdate_XOR_in_host_parent_org_counts
  ) %>%
  pivot_longer(
    c(
      CurUpdate_AND_NOT_in_host_parent_org_counts,
      CurUpdate_AND_in_host_parent_org_counts,
      CurUpdate_EQU_in_host_parent_org_counts,
      CurUpdate_NAND_in_host_parent_org_counts,
      CurUpdate_NOR_in_host_parent_org_counts,
      CurUpdate_NOT_in_host_parent_org_counts,
      CurUpdate_OR_NOT_in_host_parent_org_counts,
      CurUpdate_OR_in_host_parent_org_counts,
      CurUpdate_XOR_in_host_parent_org_counts
    ),
    names_to = "task",
    values_to = "task_in_parent_count"
  ) %>%
  mutate(
    task = str_replace(task, "CurUpdate_", "")
  ) %>%
  mutate(
    task = str_replace(task, "_in_host_parent_org_counts", "")
  ) %>%
  mutate(
    task_completed_by_parent = task_in_parent_count >= threshold,
    task = factor(
      task,
      levels = tasks,
      labels = tasks
    )
  ) %>%
  mutate(
    task_complexity = case_when(
      task == "NOT" ~ 1,
      task == "NAND" ~ 1,
      task == "OR_NOT" ~ 2,
      task == "AND" ~ 2,
      task == "OR" ~ 3,
      task == "AND_NOT" ~ 3,
      task == "NOR" ~ 4,
      task == "XOR" ~ 4,
      task == "EQU" ~ 5,
      .default = -1
    )
  )

host_parent_tasks_all_time_by_task <- host_over_time_parent_task_data %>%
  group_by(
    SEED,
    sym_condition,
    interaction_multiplier,
    task
  ) %>%
  summarize(
    n = n(),
    parent_task_completed_ever = any(task_completed_by_parent)
  ) %>%
  ungroup() %>%
  group_by(
    sym_condition,
    interaction_multiplier,
    task
  ) %>%
  summarize(
    n = n(),
    parent_tasks_completed_ever = sum(parent_task_completed_ever == TRUE)
  )
```

Next, summarize function complexity level.

```{r}
host_max_task_complexity_data <- host_over_time_parent_task_data %>%
  group_by(
    SEED,
    sym_condition,
    interaction_multiplier,
    task,
    task_complexity
  ) %>%
  summarize(
    n = n(),
    parent_task_completed_ever = any(task_completed_by_parent)
  ) %>%
  ungroup() %>%
  # This filter works fine because all replicates start with a task
  filter(
    parent_task_completed_ever
  ) %>%
  group_by(
    SEED,
    sym_condition,
    interaction_multiplier
  ) %>%
  summarize(
    max_task_complexity = max(task_complexity)
  ) %>%
  mutate(
    sym_condition = factor(
      sym_condition,
      levels = c(
        "no-syms",
        "parasite",
        "mutualist"
      )
    )
  )
```

### Host function performance - all time

```{r}
host_parent_tasks_all_time_by_task_plt <- host_parent_tasks_all_time_by_task %>%
  ggplot(
    aes(
      x = task,
      color = sym_condition,
      fill = sym_condition,
      y = parent_tasks_completed_ever
    )
  ) +
  geom_col(
    position = "dodge",
    alpha = 0.5
  ) +
  geom_text(
    aes(label = parent_tasks_completed_ever),
    position = position_dodge(width = 1),
    vjust = -0.5,
    size = 2
  ) +
  scale_color_manual(values = color_pal) +
  scale_fill_manual(values = color_pal) +
  scale_x_discrete() +
  scale_y_continuous(
    limits = c(0, 51)
  ) +
  theme(
    legend.position = "bottom",
    axis.text.x = element_text(
      angle = 90,
      hjust = 1
    )
  )
ggsave(
  file = paste(
    plot_dir,
    "host_parent_tasks_all_time_by_task_plt.pdf",
    sep = "/"
  ),
  plot = host_parent_tasks_all_time_by_task_plt,
  width = 8,
  height = 8
)
host_parent_tasks_all_time_by_task_plt
```

### Host maximum function complexity evolved

```{r}
host_max_task_complexity_plt <- host_max_task_complexity_data %>%
  ggplot(
    aes(x = sym_condition, fill = sym_condition, y = max_task_complexity)
  ) +
  geom_point(
    mapping=aes(color = sym_condition),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_x_discrete() +
  scale_color_manual(values = color_pal) +
  scale_fill_manual(values = color_pal) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(
      angle = 60,
      hjust = 1
    )
  )
ggsave(
  file = paste(plot_dir, "host_max_task_complexity_plt.pdf", sep = "/"),
  plot = host_max_task_complexity_plt,
  width = 4,
  height = 4
)
host_max_task_complexity_plt
```

Kruskal-Wallis test to check significance among distribution differences.

```{r}
kw_test <- host_max_task_complexity_data %>%
  group_by(interaction_multiplier) %>%
  rstatix::kruskal_test(max_task_complexity ~ sym_condition)
kw_test
```

Kruskal-Wallis test is significant, so we run Bonferroni-corrected pairwise Wilcoxon rank-sum tests to determine significant differences between pairs of treatments.

```{r}
pairwise.wilcox.test(
  x = host_max_task_complexity_data$max_task_complexity,
  g = host_max_task_complexity_data$sym_condition,
  p.adjust.method = "bonferroni",
  exact = FALSE
)
```

## Symbiont function performance

```{r}
# Pivot task data long, selecting only relevant columns
sym_over_time_parent_task_data <- time_data %>%
  select(
    SEED,
    sym_condition,
    interaction_multiplier,
    update,
    CurUpdate_AND_NOT_in_sym_parent_org_counts,
    CurUpdate_AND_in_sym_parent_org_counts,
    CurUpdate_EQU_in_sym_parent_org_counts,
    CurUpdate_NAND_in_sym_parent_org_counts,
    CurUpdate_NOR_in_sym_parent_org_counts,
    CurUpdate_NOT_in_sym_parent_org_counts,
    CurUpdate_OR_NOT_in_sym_parent_org_counts,
    CurUpdate_OR_in_sym_parent_org_counts,
    CurUpdate_XOR_in_sym_parent_org_counts
  ) %>%
  pivot_longer(
    c(
      CurUpdate_AND_NOT_in_sym_parent_org_counts,
      CurUpdate_AND_in_sym_parent_org_counts,
      CurUpdate_EQU_in_sym_parent_org_counts,
      CurUpdate_NAND_in_sym_parent_org_counts,
      CurUpdate_NOR_in_sym_parent_org_counts,
      CurUpdate_NOT_in_sym_parent_org_counts,
      CurUpdate_OR_NOT_in_sym_parent_org_counts,
      CurUpdate_OR_in_sym_parent_org_counts,
      CurUpdate_XOR_in_sym_parent_org_counts
    ),
    names_to = "task",
    values_to = "task_in_parent_count"
  ) %>%
  mutate(
    task = str_replace(task, "CurUpdate_", "")
  ) %>%
  mutate(
    task = str_replace(task, "_in_sym_parent_org_counts", "")
  ) %>%
  mutate(
    task_completed_by_parent = task_in_parent_count >= threshold,
    task = factor(
      task,
      levels = tasks,
      labels = tasks
    )
  ) %>%
  mutate(
    task_complexity = case_when(
      task == "NOT" ~ 1,
      task == "NAND" ~ 1,
      task == "OR_NOT" ~ 2,
      task == "AND" ~ 2,
      task == "OR" ~ 3,
      task == "AND_NOT" ~ 3,
      task == "NOR" ~ 4,
      task == "XOR" ~ 4,
      task == "EQU" ~ 5,
      .default = -1
    )
  )

sym_parent_tasks_all_time_by_task <- sym_over_time_parent_task_data %>%
  group_by(
    SEED,
    sym_condition,
    interaction_multiplier,
    task
  ) %>%
  summarize(
    n = n(),
    parent_task_completed_ever = any(task_completed_by_parent)
  ) %>%
  ungroup() %>%
  group_by(
    sym_condition,
    interaction_multiplier,
    task
  ) %>%
  summarize(
    n = n(),
    parent_tasks_completed_ever = sum(parent_task_completed_ever == TRUE)
  )
```

Next, summarize function complexity level.

```{r}
sym_max_task_complexity_data <- sym_over_time_parent_task_data %>%
  group_by(
    SEED,
    sym_condition,
    interaction_multiplier,
    task,
    task_complexity
  ) %>%
  summarize(
    n = n(),
    parent_task_completed_ever = any(task_completed_by_parent)
  ) %>%
  ungroup() %>%
  # This filter works fine because all replicates start with a task
  filter(
    parent_task_completed_ever
  ) %>%
  group_by(
    SEED,
    sym_condition,
    interaction_multiplier
  ) %>%
  summarize(
    max_task_complexity = max(task_complexity)
  ) %>%
  mutate(
    sym_condition = factor(
      sym_condition,
      levels = c(
        "no-syms",
        "parasite",
        "mutualist"
      )
    )
  )
```

### Symbiont function performance - all time

```{r}
sym_parent_tasks_all_time_by_task_plt <- sym_parent_tasks_all_time_by_task %>%
  ggplot(
    aes(
      x = task,
      color = sym_condition,
      fill = sym_condition,
      y = parent_tasks_completed_ever
    )
  ) +
  geom_col(
    position = "dodge",
    alpha = 0.5
  ) +
  geom_text(
    aes(label = parent_tasks_completed_ever),
    position = position_dodge(width = 1),
    vjust = -0.5,
    size = 2
  ) +
  scale_color_manual(values = color_pal) +
  scale_fill_manual(values = color_pal) +
  scale_x_discrete() +
  scale_y_continuous(
    limits = c(0, 51)
  ) +
  theme(
    legend.position = "bottom",
    axis.text.x = element_text(
      angle = 90,
      hjust = 1
    )
  )
ggsave(
  file = paste(plot_dir, "sym_parent_tasks_all_time_by_task_plt.pdf", sep = "/"),
  plot = sym_parent_tasks_all_time_by_task_plt,
  width = 8,
  height = 8
)
sym_parent_tasks_all_time_by_task_plt
```

### Symbiont maximum function complexity evolved

```{r}
sym_max_task_complexity_plt <- sym_max_task_complexity_data %>%
  ggplot(
    aes(x = sym_condition, fill = sym_condition, y = max_task_complexity)
  ) +
  geom_point(
    mapping=aes(color = sym_condition),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_x_discrete() +
  scale_color_manual(values = c(color_pal[2], color_pal[3])) +
  scale_fill_manual(values = c(color_pal[2], color_pal[3])) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(
      angle = 60,
      hjust = 1
    )
  )
ggsave(
  file = paste(plot_dir, "sym_max_task_complexity_plt.pdf", sep = "/"),
  plot = sym_max_task_complexity_plt,
  width = 4,
  height = 4
)
sym_max_task_complexity_plt
```

## Host phenotypic diversity

### Number of distinct function profiles

Over time:

```{r}
host_parent_num_task_sets_plt <- time_data %>%
  # sample to make plotting faster
  filter(update %% 1000 == 0) %>%
  ggplot(
    aes(
      x = update,
      y = CurUpdate_host_parent_num_task_sets,
      color = sym_condition,
      fill = sym_condition
    )
  ) +
  stat_summary(fun = "mean", geom = "line") +
  stat_summary(
    fun.data = "mean_cl_boot",
    fun.args = list(conf.int = 0.95),
    geom = "ribbon",
    alpha = 0.2,
    linetype = 0
  ) +
  scale_color_manual(values = color_pal) +
  scale_fill_manual(values = color_pal) +
  theme(
    legend.position = "bottom"
  )
ggsave(
  file = paste(plot_dir, "host_parent_num_task_sets_plt.pdf", sep = "/"),
  plot = host_parent_num_task_sets_plt,
  width = 8,
  height = 8
)
host_parent_num_task_sets_plt
```

At the end of the experiment:

```{r}
host_final_num_task_sets_elapsed_plt <- summary_data %>%
  ggplot(
    aes(
      x = sym_condition,
      fill = sym_condition,
      y = CurUpdate_host_parent_num_task_sets
    )
  ) +
  geom_point(
    mapping=aes(color = sym_condition),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_x_discrete() +
  scale_color_manual(values = color_pal) +
  scale_fill_manual(values = color_pal) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(
      angle = 60,
      hjust = 1
    )
  )
ggsave(
  file = paste(plot_dir, "host_final_num_task_sets_elapsed_plt.pdf", sep = "/"),
  plot = host_final_num_task_sets_elapsed_plt,
  width = 4,
  height = 4
)
host_final_num_task_sets_elapsed_plt
```

### Shannon diversity of function profiles

Over time:

```{r}
host_parent_entropy_task_sets_plt <- time_data %>%
  filter(update %% 1000 == 0) %>%
  ggplot(
    aes(
      x = update,
      y = CurUpdate_host_parent_entropy_task_sets,
      color = sym_condition,
      fill = sym_condition
    )
  ) +
  stat_summary(fun = "mean", geom = "line") +
  stat_summary(
    fun.data = "mean_cl_boot",
    fun.args = list(conf.int = 0.95),
    geom = "ribbon",
    alpha = 0.2,
    linetype = 0
  ) +
  scale_color_manual(values = color_pal) +
  scale_fill_manual(values = color_pal) +
  theme(
    legend.position = "bottom"
  )
ggsave(
  file = paste(plot_dir, "host_parent_entropy_task_sets_plt.pdf", sep = "/"),
  plot = host_parent_entropy_task_sets_plt,
  width = 8,
  height = 8
)
host_parent_entropy_task_sets_plt
```

At the end of the experiment:

```{r}
host_final_entopy_task_sets_elapsed_plt <- summary_data %>%
  ggplot(
    aes(
      x = sym_condition,
      fill = sym_condition,
      y = CurUpdate_host_parent_entropy_task_sets
    )
  ) +
  geom_point(
    mapping=aes(color = sym_condition),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_x_discrete() +
  theme(
    legend.position = "none",
    axis.text.x = element_text(
      angle = 60,
      hjust = 1
    )
  )
ggsave(
  file = paste(plot_dir, "host_final_entopy_task_sets_elapsed_plt.pdf", sep = "/"),
  plot = host_final_entopy_task_sets_elapsed_plt,
  width = 10,
  height = 10
)
```

Statistical analysis

```{r}
kw_test <- summary_data %>%
  group_by(interaction_multiplier) %>%
  rstatix::kruskal_test(CurUpdate_host_parent_entropy_task_sets ~ sym_condition)
kw_test
```

```{r}
pairwise.wilcox.test(
  x = summary_data$CurUpdate_host_parent_entropy_task_sets,
  g = summary_data$sym_condition,
  p.adjust.method = "bonferroni",
  exact = FALSE
)
```

## Symbiont phenotypic diversity

### Number of distinct function profiles

Over time:

```{r}
sym_parent_num_task_sets_plt <- time_data %>%
  # sample to make plotting faster
  filter(update %% 1000 == 0) %>%
  ggplot(
    aes(
      x = update,
      y = CurUpdate_sym_parent_num_task_sets,
      color = sym_condition,
      fill = sym_condition
    )
  ) +
  stat_summary(fun = "mean", geom = "line") +
  stat_summary(
    fun.data = "mean_cl_boot",
    fun.args = list(conf.int = 0.95),
    geom = "ribbon",
    alpha = 0.2,
    linetype = 0
  ) +
  scale_color_manual(values = color_pal) +
  scale_fill_manual(values = color_pal) +
  theme(
    legend.position = "bottom"
  )
ggsave(
  file = paste(plot_dir, "sym_parent_num_task_sets_plt.pdf", sep = "/"),
  plot = sym_parent_num_task_sets_plt,
  width = 8,
  height = 8
)
sym_parent_num_task_sets_plt
```

At the end of the experiment:

```{r}
sym_final_num_task_sets_elapsed_plt <- summary_data %>%
  ggplot(
    aes(
      x = sym_condition,
      fill = sym_condition,
      y = CurUpdate_sym_parent_num_task_sets
    )
  ) +
  geom_point(
    mapping=aes(color = sym_condition),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_x_discrete() +
  scale_color_manual(values = color_pal) +
  scale_fill_manual(values = color_pal) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(
      angle = 60,
      hjust = 1
    )
  )
ggsave(
  file = paste(plot_dir, "sym_final_num_task_sets_elapsed_plt.pdf", sep = "/"),
  plot = sym_final_num_task_sets_elapsed_plt,
  width = 4,
  height = 4
)
sym_final_num_task_sets_elapsed_plt
```

### Shannon diversity of function profiles

Over time:

```{r}
sym_parent_entropy_task_sets_plt <- time_data %>%
  filter(update %% 1000 == 0) %>%
  ggplot(
    aes(
      x = update,
      y = CurUpdate_sym_parent_entropy_task_sets,
      color = sym_condition,
      fill = sym_condition
    )
  ) +
  stat_summary(fun = "mean", geom = "line") +
  stat_summary(
    fun.data = "mean_cl_boot",
    fun.args = list(conf.int = 0.95),
    geom = "ribbon",
    alpha = 0.2,
    linetype = 0
  ) +
  scale_color_manual(values = color_pal) +
  scale_fill_manual(values = color_pal) +
  theme(
    legend.position = "bottom"
  )
ggsave(
  file = paste(plot_dir, "sym_parent_entropy_task_sets_plt.pdf", sep = "/"),
  plot = sym_parent_entropy_task_sets_plt,
  width = 8,
  height = 8
)
sym_parent_entropy_task_sets_plt
```

At the end of the experiment:

```{r}
sym_final_entopy_task_sets_elapsed_plt <- summary_data %>%
  ggplot(
    aes(
      x = sym_condition,
      fill = sym_condition,
      y = CurUpdate_sym_parent_entropy_task_sets
    )
  ) +
  geom_point(
    mapping=aes(color = sym_condition),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_x_discrete() +
  theme(
    legend.position = "none",
    axis.text.x = element_text(
      angle = 60,
      hjust = 1
    )
  )
ggsave(
  file = paste(plot_dir, "sym_final_entopy_task_sets_elapsed_plt.pdf", sep = "/"),
  plot = sym_final_entopy_task_sets_elapsed_plt,
  width = 10,
  height = 10
)
```

## Trait matches over time

Counts of host-symbiont pairs with any trait matches.
Note that organisms that were just born will not have had time to express any
Boolean logic functions, so they will appear as not matching.

```{r}
host_sym_any_matches_total_plt <- time_data %>%
  filter(update %% 1000 == 0) %>%
  ggplot(
    aes(
      x = update,
      y = CurUpdate_host_sym_any_matches_total,
      color = sym_condition,
      fill = sym_condition
    )
  ) +
  stat_summary(fun = "mean", geom = "line") +
  stat_summary(
    fun.data = "mean_cl_boot",
    fun.args = list(conf.int = 0.95),
    geom = "ribbon",
    alpha = 0.2,
    linetype = 0
  ) +
  scale_color_manual(values = color_pal) +
  scale_fill_manual(values = color_pal) +
  theme(
    legend.position = "bottom"
  )
ggsave(
  file = paste(plot_dir, "host_sym_any_matches_total_plt.pdf", sep = "/"),
  plot = host_sym_any_matches_total_plt,
  width = 8,
  height = 8
)
host_sym_any_matches_total_plt
```

## Evolutionary engagement

Use trait matching metrics to determine if host-symbiont populations are evolutionarily engaged.

```{r}
engagement <- summary_data %>%
  mutate(
    host_sym_engagement = CurUpdate_host_sym_any_matches_total > threshold,
  ) %>%
  group_by(
    sym_condition
  ) %>%
  summarise(
    host_sym_engagement_count = sum(host_sym_engagement)
  )
```

```{r}
host_sym_engagement_count_plt <- engagement %>%
  ggplot(
    aes(
      x = sym_condition,
      color = sym_condition,
      fill = sym_condition,
      y = host_sym_engagement_count
    )
  ) +
  geom_col(
    position = "dodge",
    alpha = 0.5
  ) +
  geom_text(
    aes(label = host_sym_engagement_count),
    position = position_dodge(width = 1),
    vjust = -0.5,
    size = 2
  ) +
  scale_color_manual(values = color_pal) +
  scale_fill_manual(values = color_pal) +
  scale_x_discrete() +
  scale_y_continuous(
    limits = c(0, 51)
  ) +
  theme(
    legend.position = "none"
  )

ggsave(
  file = paste(plot_dir, "host_sym_engagement_count_plt.pdf", sep = "/"),
  plot = host_sym_engagement_count_plt,
  width = 4,
  height = 4
)
host_sym_engagement_count_plt
```

## Generations of evolution elapsed

### Hosts

```{r}
host_avg_gens_elapsed_plt <- summary_data %>%
  ggplot(
    aes(
      x = sym_condition,
      fill = sym_condition,
      y = CurUpdate_host_mean_generations
    )
  ) +
  geom_point(
    mapping = aes(color = sym_condition),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_x_discrete() +
  scale_color_manual(values = color_pal) +
  scale_fill_manual(values = color_pal) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(
      angle = 60,
      hjust = 1
    )
  )
ggsave(
  file = paste(plot_dir, "host_avg_gens_elapsed_plt.pdf", sep = "/"),
  plot = host_avg_gens_elapsed_plt,
  width = 4,
  height = 4
)
host_avg_gens_elapsed_plt
```

```{r}
summary_data %>%
  group_by(
    sym_condition
  ) %>%
  summarise(
    mean_gens = mean(CurUpdate_host_mean_generations),
    median_gens = median(CurUpdate_host_mean_generations),
    var_gens = var(CurUpdate_host_mean_generations),
    sdev_gens = sd(CurUpdate_host_mean_generations)
  )
```

```{r}
kw_test <- summary_data %>%
  rstatix::kruskal_test(CurUpdate_host_mean_generations ~ sym_condition)
kw_test
```

```{r}
pairwise.wilcox.test(
  x = summary_data$CurUpdate_host_mean_generations,
  g = summary_data$sym_condition,
  p.adjust.method = "bonferroni",
  exact = FALSE
)
```

### Symbionts

```{r}
sym_avg_gens_elapsed_plt <- summary_data %>%
  ggplot(
    aes(
      x = sym_condition,
      fill = sym_condition,
      y = CurUpdate_sym_mean_generations
    )
  ) +
  geom_point(
    mapping = aes(color = sym_condition),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_x_discrete() +
  scale_color_manual(values = color_pal) +
  scale_fill_manual(values = color_pal) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(
      angle = 60,
      hjust = 1
    )
  )
ggsave(
  file = paste(plot_dir, "sym_avg_gens_elapsed_plt.pdf", sep = "/"),
  plot = sym_avg_gens_elapsed_plt,
  width = 4,
  height = 4
)
sym_avg_gens_elapsed_plt
```

```{r}
summary_data %>%
  group_by(
    sym_condition
  ) %>%
  summarise(
    mean_gens = mean(CurUpdate_sym_mean_generations),
    median_gens = median(CurUpdate_sym_mean_generations),
    var_gens = var(CurUpdate_sym_mean_generations),
    sdev_gens = sd(CurUpdate_sym_mean_generations)
  )
```

## Symbiont transmission rates

### Horizontal transmission attempts

```{r}
TransmissionRates_attempts_horiztrans_plt <- time_data %>%
  filter(update %% 1000 == 0) %>%
  ggplot(
    aes(
      x = update,
      y = TransmissionRates_attempts_horiztrans,
      color = sym_condition,
      fill = sym_condition
    )
  ) +
  stat_summary(fun = "mean", geom = "line") +
  stat_summary(
    fun.data = "mean_cl_boot",
    fun.args = list(conf.int = 0.95),
    geom = "ribbon",
    alpha = 0.2,
    linetype = 0
  ) +
  scale_color_manual(values = color_pal) +
  scale_fill_manual(values = color_pal) +
  theme(
    legend.position = "bottom"
  )
ggsave(
  file = paste(
    plot_dir,
    "TransmissionRates_attempts_horiztrans_plt.pdf",
    sep = "/"
  ),
  plot = TransmissionRates_attempts_horiztrans_plt,
  width = 8,
  height = 8
)
TransmissionRates_attempts_horiztrans_plt
```

### Horizontal transmission successes

```{r}
TransmissionRates_successes_horiztrans_plt <- time_data %>%
  filter(update %% 1000 == 0) %>%
  ggplot(
    aes(
      x = update,
      y = TransmissionRates_successes_horiztrans,
      color = sym_condition,
      fill = sym_condition
    )
  ) +
  stat_summary(fun = "mean", geom = "line") +
  stat_summary(
    fun.data = "mean_cl_boot",
    fun.args = list(conf.int = 0.95),
    geom = "ribbon",
    alpha = 0.2,
    linetype = 0
  ) +
  scale_color_manual(values = color_pal) +
  scale_fill_manual(values = color_pal) +
  theme(
    legend.position = "bottom"
  )
ggsave(
  file = paste(
    plot_dir,
    "TransmissionRates_successes_horiztrans_plt.pdf",
    sep = "/"
  ),
  plot = TransmissionRates_successes_horiztrans_plt,
  width = 8,
  height = 8
)
TransmissionRates_successes_horiztrans_plt
```

### Vertical transmission attempts

```{r}
TransmissionRates_attempts_verttrans_plt <- time_data %>%
  filter(update %% 1000 == 0) %>%
  ggplot(
    aes(
      x = update,
      y = TransmissionRates_attempts_verttrans,
      color = sym_condition,
      fill = sym_condition
    )
  ) +
  stat_summary(fun = "mean", geom = "line") +
  stat_summary(
    fun.data = "mean_cl_boot",
    fun.args = list(conf.int = 0.95),
    geom = "ribbon",
    alpha = 0.2,
    linetype = 0
  ) +
  scale_color_manual(values = color_pal) +
  scale_fill_manual(values = color_pal) +
  theme(
    legend.position = "bottom"
  )
ggsave(
  file = paste(
    plot_dir,
    "TransmissionRates_attempts_verttrans_plt.pdf",
    sep = "/"
  ),
  plot = TransmissionRates_attempts_verttrans_plt,
  width = 8,
  height = 8
)
TransmissionRates_attempts_verttrans_plt
```

### Successes vertical transmission

```{r}
TransmissionRates_successes_verttrans_plt <- time_data %>%
  filter(update %% 1000 == 0) %>%
  ggplot(
    aes(
      x = update,
      y = TransmissionRates_successes_verttrans,
      color = sym_condition,
      fill = sym_condition
    )
  ) +
  stat_summary(fun = "mean", geom = "line") +
  stat_summary(
    fun.data = "mean_cl_boot",
    fun.args = list(conf.int = 0.95),
    geom = "ribbon",
    alpha = 0.2,
    linetype = 0
  ) +
  scale_color_manual(values = color_pal) +
  scale_fill_manual(values = color_pal) +
  theme(
    legend.position = "bottom"
  )
ggsave(
  file = paste(
    plot_dir,
    "TransmissionRates_successes_verttrans_plt.pdf",
    sep = "/"
  ),
  plot = TransmissionRates_successes_verttrans_plt,
  width = 8,
  height = 8
)
TransmissionRates_successes_verttrans_plt
```

## Population size

```{r}
pop_size_summary <- time_data %>%
  group_by(
    SEED,
    sym_condition
  ) %>%
  summarize(
    host_pop_size_avg = mean(OrgCounts_host_count),
    host_pop_size_median = median(OrgCounts_host_count),
    host_pop_size_variance = var(OrgCounts_host_count),
    sym_pop_size_avg = mean(OrgCounts_hosted_sym_count),
    sym_pop_size_median = median(OrgCounts_hosted_sym_count),
    sym_pop_size_variance = var(OrgCounts_hosted_sym_count)
  )
```

### Host population size

```{r}
host_pop_size_ot_plt <- time_data %>%
  filter(update %% 1000 == 0) %>%
  ggplot(
    aes(
      x = update,
      y = OrgCounts_host_count,
      color = sym_condition,
      fill = sym_condition
    )
  ) +
  stat_summary(fun = "mean", geom = "line") +
  stat_summary(
    fun.data = "mean_cl_boot",
    fun.args = list(conf.int = 0.95),
    geom = "ribbon",
    alpha = 0.2,
    linetype = 0
  ) +
  scale_color_manual(values = color_pal) +
  scale_fill_manual(values = color_pal) +
  theme(
    legend.position = "bottom"
  )
ggsave(
  file = paste(
    plot_dir,
    "host_pop_size_ot_plt.pdf",
    sep = "/"
  ),
  plot = host_pop_size_ot_plt,
  width = 8,
  height = 8
)
host_pop_size_ot_plt
```

### Symbiont population size

```{r}
sym_pop_size_ot_plt <- time_data %>%
  filter(update %% 1000 == 0) %>%
  ggplot(
    aes(
      x = update,
      y = OrgCounts_hosted_sym_count,
      color = sym_condition,
      fill = sym_condition
    )
  ) +
  stat_summary(fun = "mean", geom = "line") +
  stat_summary(
    fun.data = "mean_cl_boot",
    fun.args = list(conf.int = 0.95),
    geom = "ribbon",
    alpha = 0.2,
    linetype = 0
  ) +
  scale_color_manual(values = color_pal) +
  scale_fill_manual(values = color_pal) +
  theme(
    legend.position = "bottom"
  )
ggsave(
  file = paste(
    plot_dir,
    "sym_pop_size_ot_plt.pdf",
    sep = "/"
  ),
  plot = sym_pop_size_ot_plt,
  width = 8,
  height = 8
)
sym_pop_size_ot_plt
```

## Manuscript figures

```{r}
host_max_task_complexity_data_ci <- host_max_task_complexity_data %>%
  group_by(
    sym_condition
  ) %>%
  summarize(
    mean_max_task_complexity = mean(max_task_complexity),
    min_max_task_complexity = min(max_task_complexity),
    max_max_task_complexity = max(max_task_complexity),
    ci = list(
      mean_cl_boot(
        max_task_complexity,
        conf.int = 0.99,
        B = 10000,
      ) %>% rename(mean = y, lwr = ymin, upr = ymax)
    )
  ) %>%
  unnest(cols = c(ci)) %>%
  mutate(
    sym_condition = factor(
      sym_condition,
      levels = c("no-syms", "parasite", "mutualist"),
      labels = c("No symbionts", "Parasites", "Mutualists")
    )
  )

dist_data <- host_max_task_complexity_data %>%
  mutate(
    sym_condition = factor(
      sym_condition,
      levels = c("no-syms", "parasite", "mutualist"),
      labels = c("No symbionts", "Parasites", "Mutualists")
    )
  )

manuscript_host_task_complexity_plt <- host_max_task_complexity_data_ci %>%
  ggplot(
    aes(
      x = sym_condition,
      fill = sym_condition,
      y = mean_max_task_complexity
    )
  ) +
  geom_bar(stat = "identity") +
  geom_point(
    data = dist_data,
    mapping = aes(x = sym_condition, y = max_task_complexity),
    position = position_jitter(width = .4, height = 0.05),
    size = .7,
    alpha = 0.9,
    color = "darkgray",
    fill = "darkgray"
  ) +
  geom_linerange(
    aes(
      ymin = lwr,
      ymax = upr
    ),
    alpha = 0.9,
    linewidth = 0.7,
    color = "black"
  ) +
  scale_x_discrete(
    limits = c("No symbionts", "Parasites", "Mutualists"),
    labels = c("None", "Par.", "Mut.")
  ) +
  scale_fill_manual(values = color_pal) +
  labs(
    x = "Sym. type",
    y = "Host function complexity"
  ) +
  theme(
    legend.position = "none",
    axis.title.y = element_text(size = 10),
    axis.text.y = element_text(size = 8),
    axis.title.x = element_text(size = 10),
    axis.text.x = element_text(size = 8)
  )


ggsave(
  file = paste(
    plot_dir,
    "manuscript_flat_rewards_host_task_complexity_plt.pdf",
    sep = "/"
  ),
  plot = manuscript_host_task_complexity_plt,
  width = 3,
  height = 3
)
```