# 2026-02-23 - Task credit constraint

## Dependencies and setup

```{r}
library(tidyverse)
library(cowplot)
library(RColorBrewer)
library(khroma)
library(rstatix)
library(knitr)
library(kableExtra)
library(viridis)
library(boot)
```

```{r}
# Build relative path from R's working directory to what is actually the working
# directory for this experiment.
experiment_slug <- "2026-02-23-health-task-credit"
working_directory <- paste(
  "experiments",
  experiment_slug,
  "analysis",
  sep = "/"
)
```

```{r}
# Use the cowplot theme as our default for ggplots
theme_set(theme_cowplot())
# Create directory to save plots
plot_dir <- paste(
  working_directory,
  "plots",
  sep = "/"
)
dir.create(
  plot_dir,
  showWarnings = FALSE
)
```

### Load and preprocess data

```{r}
tasks <- c("NOT", "NAND", "OR_NOT", "AND", "OR", "AND_NOT", "NOR", "XOR", "EQU")
max_pop_size <- 100 * 100
threshold <- 0.025 * max_pop_size
num_replicates <- 50
```

Load summary data.

```{r}
summary_data_path <- paste(
  working_directory,
  "data",
  "summary.csv",
  sep = "/"
)
summary_data <- read_csv(summary_data_path)
# Do any universal preprocessing
summary_data <- summary_data %>%
  mutate(
    SEED = as.factor(SEED),
    START_MOI = as.factor(START_MOI),
    VERTICAL_TRANSMISSION = as.factor(VERTICAL_TRANSMISSION),
    HEALTH_TYPE = as.factor(HEALTH_TYPE),
    HEALTH_INTERACTION_CHANCE = as.factor(HEALTH_INTERACTION_CHANCE),
    MUTUALIST_CYCLE_GAIN_PROP = as.factor(MUTUALIST_CYCLE_GAIN_PROP),
    PARASITE_CYCLE_LOSS_PROP = as.factor(PARASITE_CYCLE_LOSS_PROP),
    PARASITE_BASE_CYCLE_PROP = as.factor(PARASITE_BASE_CYCLE_PROP),
    MUTUALIST_CYCLE_DONATE_MULTIPLIER = as.factor(MUTUALIST_CYCLE_DONATE_MULTIPLIER), # nolint: line_length_linter.
    PARASITE_CYCLE_STEAL_MULTIPLIER = as.factor(PARASITE_CYCLE_STEAL_MULTIPLIER), # nolint: line_length_linter.
    HOST_ONLY_FIRST_TASK_CREDIT = as.factor(HOST_ONLY_FIRST_TASK_CREDIT),
    SYM_ONLY_FIRST_TASK_CREDIT = as.factor(SYM_ONLY_FIRST_TASK_CREDIT),
    SGP_MUT_PER_BIT_RATE = as.factor(SGP_MUT_PER_BIT_RATE)
  ) %>%
  mutate(
    task_credit = paste0(
      "H", HOST_ONLY_FIRST_TASK_CREDIT,
      "S", SYM_ONLY_FIRST_TASK_CREDIT
    ),
    sym_condition = case_when(
      START_MOI == "0" ~ "no-syms",
      START_MOI == "1" & HEALTH_TYPE == "parasite" ~ "parasite",
      START_MOI == "1" & HEALTH_TYPE == "mutualist" ~ "mutualist",
      .default = "unknown" # precaution against unexpected parameterizations
    ),
    interaction_multiplier = as.factor(PARASITE_CYCLE_STEAL_MULTIPLIER),
    extinct_syms = OrgCounts_hosted_sym_count == 0
  ) %>%
  mutate(
    treatment = paste0(
      sym_condition,
      "_", task_credit
    ),
    task_credit = as.factor(task_credit)
  ) %>%
  mutate(
    treatment = as.factor(
      treatment
    )
  )
```

Load time series data.

```{r}
time_data_path <- paste(
  working_directory,
  "data",
  "time_series.csv",
  sep = "/"
)
time_data <- read_csv(time_data_path)

# Do any universal preprocessing
time_data <- time_data %>%
  mutate(
    SEED = as.factor(SEED),
    START_MOI = as.factor(START_MOI),
    VERTICAL_TRANSMISSION = as.factor(VERTICAL_TRANSMISSION),
    HEALTH_TYPE = as.factor(HEALTH_TYPE),
    HOST_ONLY_FIRST_TASK_CREDIT = as.factor(HOST_ONLY_FIRST_TASK_CREDIT),
    SYM_ONLY_FIRST_TASK_CREDIT = as.factor(SYM_ONLY_FIRST_TASK_CREDIT)
  ) %>%
  mutate(
    sym_condition = case_when(
      START_MOI == "0" ~ "no-syms",
      START_MOI == "1" & HEALTH_TYPE == "parasite" ~ "parasite",
      START_MOI == "1" & HEALTH_TYPE == "mutualist" ~ "mutualist",
      .default = "unknown" # precaution against unexpected parameterizations
    ),
    task_credit = paste0(
      "H", HOST_ONLY_FIRST_TASK_CREDIT,
      "S", SYM_ONLY_FIRST_TASK_CREDIT
    )
  ) %>%
  mutate(
    treatment = paste0(
      sym_condition,
      "_", task_credit
    ),
    task_credit = as.factor(task_credit)
  ) %>%
  mutate(
    treatment = as.factor(
      treatment
    )
  )
```

```{r}
# Pivot task data long:
# 1) Select only relevant columns
host_over_time_parent_task_data <- time_data %>%
  select(
    SEED,
    treatment,
    sym_condition,
    task_credit,
    update,
    CurUpdate_AND_NOT_in_host_parent_org_counts,
    CurUpdate_AND_in_host_parent_org_counts,
    CurUpdate_EQU_in_host_parent_org_counts,
    CurUpdate_NAND_in_host_parent_org_counts,
    CurUpdate_NOR_in_host_parent_org_counts,
    CurUpdate_NOT_in_host_parent_org_counts,
    CurUpdate_OR_NOT_in_host_parent_org_counts,
    CurUpdate_OR_in_host_parent_org_counts,
    CurUpdate_XOR_in_host_parent_org_counts
  ) %>%
  pivot_longer(
    c(
      CurUpdate_AND_NOT_in_host_parent_org_counts,
      CurUpdate_AND_in_host_parent_org_counts,
      CurUpdate_EQU_in_host_parent_org_counts,
      CurUpdate_NAND_in_host_parent_org_counts,
      CurUpdate_NOR_in_host_parent_org_counts,
      CurUpdate_NOT_in_host_parent_org_counts,
      CurUpdate_OR_NOT_in_host_parent_org_counts,
      CurUpdate_OR_in_host_parent_org_counts,
      CurUpdate_XOR_in_host_parent_org_counts
    ),
    names_to = "task",
    values_to = "task_in_parent_count"
  ) %>%
  mutate(
    task = str_replace(task, "CurUpdate_", "")
  ) %>%
  mutate(
    task = str_replace(task, "_in_host_parent_org_counts", "")
  ) %>%
  mutate(
    task_completed_by_parent = task_in_parent_count >= threshold,
    task = factor(
      task,
      levels = tasks,
      labels = tasks
    )
  ) %>%
  mutate(
    task_complexity = case_when(
      task == "NOT" ~ 1,
      task == "NAND" ~ 1,
      task == "OR_NOT" ~ 2,
      task == "AND" ~ 2,
      task == "OR" ~ 3,
      task == "AND_NOT" ~ 3,
      task == "NOR" ~ 4,
      task == "XOR" ~ 4,
      task == "EQU" ~ 5,
      .default = -1
    )
  )

host_parent_tasks_all_time_by_task <- host_over_time_parent_task_data %>%
  group_by(
    SEED,
    treatment,
    sym_condition,
    task_credit,
    task
  ) %>%
  dplyr::summarize(
    n = n(),
    parent_task_completed_ever = any(task_completed_by_parent)
  ) %>%
  ungroup() %>%
  group_by(
    treatment,
    sym_condition,
    task_credit,
    task
  ) %>%
  dplyr::summarize(
    n = n(),
    parent_tasks_completed_ever = sum(parent_task_completed_ever == TRUE)
  )
```

Additionally, summarize by task complexity level.

```{r}
host_max_task_complexity_data <- host_over_time_parent_task_data %>%
  group_by(
    SEED,
    treatment,
    sym_condition,
    task_credit,
    task,
    task_complexity
  ) %>%
  summarise(
    n = n(),
    parent_task_completed_ever = any(task_completed_by_parent)
  ) %>%
  ungroup() %>%
  # This filter works fine because all replicates start with a task
  filter(
    parent_task_completed_ever
  ) %>%
  group_by(
    SEED,
    treatment,
    sym_condition,
    task_credit
  ) %>%
  summarise(
    max_task_complexity = max(task_complexity)
  ) %>%
  mutate(
    sym_condition = factor(
      sym_condition,
      levels = c(
        "no-syms",
        "parasite",
        "mutualist"
      )
    )
  ) %>%
  mutate(
    experiment = "2026-02-23-health-task-credit"
  )
```

```{r}
write_csv(
  host_max_task_complexity_data,
  paste(
    working_directory,
    "data",
    "host_max_task_complexity.csv",
    sep = "/"
  )
)
```

## Calculate baseline for constraint

Use all no-symbiont replicates.

Calculate a bootstrapped 99% confidence interval for mean function complexity achieved
in the absence of symbionts.

```{r}
# What is the 95% cutoff for number of replicates?
cut_off_95 <- ceil(num_replicates * 0.95)
no_sym_max_task_complexity_thresholds <- host_max_task_complexity_data %>%
  ungroup() %>%
  filter(sym_condition == "no-syms") %>%
  group_by(
    task_credit
  ) %>%
  arrange(
    desc(max_task_complexity)
  ) %>%
  summarise(
    complexity_threshold = nth(max_task_complexity, n = cut_off_95),
    median_complexity = median(max_task_complexity),
    min_complexity = min(max_task_complexity),
    max_complexity = max(max_task_complexity)
  )
```

Next, categorize each replicate of the mutualism condition (across synergy levels) as constraining, promoting, or neither.

```{r}
mutualism_effect_on_hosts <- host_max_task_complexity_data %>%
  ungroup() %>%
  filter(sym_condition == "mutualist") %>%
  left_join(no_sym_max_task_complexity_thresholds, by = c("task_credit")) %>%
  mutate(
    constrains = max_task_complexity < complexity_threshold
  ) %>%
  mutate(
    mutualist_effect = case_when(
      constrains ~ "constraint",
      .default = "unconstrained"
    )
  ) %>%
  group_by(
    task_credit,
    sym_condition
  ) %>%
  summarise(
    n = n(),
    constrains_count = sum(constrains),
    no_constraint_count = sum(!constrains)
  )
```

## Plot effects across synergy levels

```{r}
constraint_plt <- mutualism_effect_on_hosts %>%
  ggplot(
    aes(
      x = task_credit,
      color = task_credit,
      fill = task_credit,
      y = constrains_count
    )
  ) +
  geom_col(
    position = "dodge",
    alpha = 0.5
  ) +
  geom_text(
    aes(label = constrains_count),
    position = position_dodge(width = 1),
    vjust = -0.5,
    size = 2
  ) +
  scale_color_dark() +
  scale_fill_dark() +
  scale_x_discrete() +
  scale_y_continuous(
    limits = c(0, 51)
  ) +
  theme(
    legend.position = "none"
  )

ggsave(
  file = paste(plot_dir, "task_credit_constraint_plt.pdf", sep = "/"),
  plot = constraint_plt,
  width = 4,
  height = 4
)
```

## Max task complexity

```{r}

host_max_task_complexity_data_ci <- host_max_task_complexity_data %>%
  group_by(
    treatment,
    sym_condition,
    task_credit
  ) %>%
  summarise(
    mean_max_task_complexity = mean(max_task_complexity),
    min_max_task_complexity = min(max_task_complexity),
    max_max_task_complexity = max(max_task_complexity),
    ci = list(
      mean_cl_boot(
        max_task_complexity,
        conf.int = 0.99,
        B = 10000,
      ) %>% rename(mean = y, lwr = ymin, upr = ymax)
    )
  ) %>%
  unnest(cols = c(ci))
  # %>%
  # mutate(
  #   sym_condition = factor(
  #     sym_condition,
  #     levels = c("no-syms", "parasite", "mutualist"),
  #     labels = c("No sym.", "Par.", "Mut.")
  #   )
  # )

host_max_complexity_bar_plt <- host_max_task_complexity_data_ci %>%
  ggplot(
    aes(
      x = sym_condition,
      fill = sym_condition,
      y = mean
    )
  ) +
  geom_bar(stat = "identity") +
  geom_point(
    data = host_max_task_complexity_data,
    aes(x = sym_condition, y = max_task_complexity),
    position = position_jitter(width = .3, height = 0.1),
    size = .7,
    alpha = 0.6,
    color = "darkgray",
    fill = "darkgray"
  ) +
  geom_linerange(
    aes(
      ymin = lwr,
      ymax = upr
    ),
    alpha = 0.9,
    linewidth = 0.5
  ) +
  scale_x_discrete() +
  scale_fill_bright() +
  scale_y_continuous(
    limits = c(0, 5.1)
  ) +
  labs(
    x = "Symbiont type",
    y = "Mean function complexity"
  ) +
  facet_wrap(~task_credit, nrow = 1) +
  theme(
    legend.position = "none"
    # axis.text.x = element_text(
    #   angle = 45,
    #   hjust = 1
    # )
  )
ggsave(
  file = paste(plot_dir, "host_max_complexity_bar_plt.pdf", sep = "/"),
  plot = host_max_complexity_bar_plt,
  width = 8,
  height = 4
)
```

## Engagement

```{r}
engagement <- summary_data %>%
  mutate(
    host_sym_engagement = CurUpdate_host_sym_any_matches_total > threshold,
  ) %>%
  group_by(
    task_credit,
    sym_condition
  ) %>%
  summarise(
    host_sym_engagement_count = sum(host_sym_engagement)
  )

# filter(summary_data, sym_condition == "mutualist") %>%
#   select(task_credit, sym_condition, CurUpdate_host_sym_any_matches_total)
```

H0 treatments are not constrained, but are engaged
because hosts can evolve complex tasks without giving up simple tasks (and compatibility with mutualist).

```{r}
host_sym_engagement_count_plt <- engagement %>%
  filter(sym_condition == "mutualist") %>%
  ggplot(
    aes(
      x = task_credit,
      color = task_credit,
      fill = task_credit,
      y = host_sym_engagement_count
    )
  ) +
  geom_col(
    position = "dodge",
    alpha = 0.5
  ) +
  geom_text(
    aes(label = host_sym_engagement_count),
    position = position_dodge(width = 1),
    vjust = -0.5,
    size = 2
  ) +
  scale_color_dark() +
  scale_fill_dark() +
  scale_x_discrete() +
  scale_y_continuous(
    limits = c(0, 51)
  ) +
  theme(
    legend.position = "none"
  )

ggsave(
  file = paste(plot_dir, "host_sym_engagement_count_plt.pdf", sep = "/"),
  plot = host_sym_engagement_count_plt,
  width = 4,
  height = 4
)
```