# 2025-12-09 - Comparing mutualism constraint across function reward levels

## Dependencies and setup

```{r}
library(tidyverse)
library(cowplot)
library(RColorBrewer)
library(khroma)
library(rstatix)
library(knitr)
library(kableExtra)
library(viridis)
# library(webshot2)
```

```{r}
# Build relative path from R's working directory to what is actually the working
# directory for this experiment.
experiment_slug <- "2025-12-09-health-big-rewards"
working_directory <- paste(
  "experiments",
  experiment_slug,
  "analysis",
  sep = "/"
)
```

```{r}
# Use the cowplot theme as our default for ggplots
theme_set(theme_cowplot())
# Create directory to save plots
plot_dir <- paste(
  working_directory,
  "plots",
  sep = "/"
)
dir.create(
  plot_dir,
  showWarnings = FALSE
)
```

### Load and preprocess data

```{r}
tasks <- c("NOT", "NAND", "OR_NOT", "AND", "OR", "AND_NOT", "NOR", "XOR", "EQU")
max_pop_size <- 100 * 100
threshold <- 0.025 * max_pop_size
```

Load summary data.

```{r}
summary_data_path <- paste(
  working_directory,
  "data",
  "summary.csv",
  sep = "/"
)
summary_data <- read_csv(summary_data_path)
# Do any universal preprocessing
summary_data <- summary_data %>%
  mutate(
    SEED = as.factor(SEED),
    START_MOI = as.factor(START_MOI),
    VERTICAL_TRANSMISSION = as.factor(VERTICAL_TRANSMISSION),
    HEALTH_TYPE = as.factor(HEALTH_TYPE),
    HEALTH_INTERACTION_CHANCE = as.factor(HEALTH_INTERACTION_CHANCE),
    MUTUALIST_CYCLE_GAIN_PROP = as.factor(MUTUALIST_CYCLE_GAIN_PROP),
    PARASITE_CYCLE_LOSS_PROP = as.factor(PARASITE_CYCLE_LOSS_PROP),
    PARASITE_BASE_CYCLE_PROP = as.factor(PARASITE_BASE_CYCLE_PROP),
    MUTUALIST_CYCLE_DONATE_MULTIPLIER = as.factor(MUTUALIST_CYCLE_DONATE_MULTIPLIER), # nolint: line_length_linter.
    PARASITE_CYCLE_STEAL_MULTIPLIER = as.factor(PARASITE_CYCLE_STEAL_MULTIPLIER), # nolint: line_length_linter.
    HOST_ONLY_FIRST_TASK_CREDIT = as.factor(HOST_ONLY_FIRST_TASK_CREDIT),
    SYM_ONLY_FIRST_TASK_CREDIT = as.factor(SYM_ONLY_FIRST_TASK_CREDIT),
    TASK_ENV_CFG_PATH = as.factor(TASK_ENV_CFG_PATH)
  ) %>%
  mutate(
    task_credit = paste0(
      "H", HOST_ONLY_FIRST_TASK_CREDIT,
      "S", SYM_ONLY_FIRST_TASK_CREDIT
    ),
    sym_condition = case_when(
      START_MOI == "0" ~ "no-syms",
      START_MOI == "1" & HEALTH_TYPE == "parasite" ~ "parasite",
      START_MOI == "1" & HEALTH_TYPE == "mutualist" ~ "mutualist",
      .default = "unknown" # precaution against unexpected parameterizations
    ),
    interaction_multiplier = as.factor(PARASITE_CYCLE_STEAL_MULTIPLIER),
    environment = case_when(
      TASK_ENV_CFG_PATH == "environment-rewards-2.json" ~ "x2",
      TASK_ENV_CFG_PATH == "environment-rewards-4.json" ~ "x4",
      TASK_ENV_CFG_PATH == "environment-rewards-8.json" ~ "x8",
      TASK_ENV_CFG_PATH == "environment-rewards-16.json" ~ "x16",
      .default = "unknown"
    )
  ) %>%
  mutate(
    treatment = paste0(
      sym_condition,
      "_E=", environment,
      "_IM=", interaction_multiplier
    ),
    environment = factor(
      environment,
      levels = c("x2", "x4", "x8", "x16")
    )
  ) %>%
  mutate(
    treatment = as.factor(
      treatment
    )
  )
```

No extinctions

Load time series data.

```{r}
time_data_path <- paste(
  working_directory,
  "data",
  "time_series.csv",
  sep = "/"
)
time_data <- read_csv(time_data_path)

# Do any universal preprocessing
time_data <- time_data %>%
  mutate(
    SEED = as.factor(SEED),
    START_MOI = as.factor(START_MOI),
    VERTICAL_TRANSMISSION = as.factor(VERTICAL_TRANSMISSION),
    HEALTH_TYPE = as.factor(HEALTH_TYPE),
    MUTUALIST_CYCLE_DONATE_MULTIPLIER = as.factor(MUTUALIST_CYCLE_DONATE_MULTIPLIER), # nolint: line_length_linter.
    PARASITE_CYCLE_STEAL_MULTIPLIER = as.factor(PARASITE_CYCLE_STEAL_MULTIPLIER), # nolint: line_length_linter.
    TASK_ENV_CFG_PATH = as.factor(TASK_ENV_CFG_PATH)
  ) %>%
  mutate(
    sym_condition = case_when(
      START_MOI == "0" ~ "no-syms",
      START_MOI == "1" & HEALTH_TYPE == "parasite" ~ "parasite",
      START_MOI == "1" & HEALTH_TYPE == "mutualist" ~ "mutualist",
      .default = "unknown" # precaution against unexpected parameterizations
    ),
    interaction_multiplier = as.factor(PARASITE_CYCLE_STEAL_MULTIPLIER),
    environment = case_when(
      TASK_ENV_CFG_PATH == "environment-rewards-2.json" ~ "x2",
      TASK_ENV_CFG_PATH == "environment-rewards-4.json" ~ "x4",
      TASK_ENV_CFG_PATH == "environment-rewards-8.json" ~ "x8",
      TASK_ENV_CFG_PATH == "environment-rewards-16.json" ~ "x16",
      .default = "unknown"
    )
  ) %>%
  mutate(
    treatment = paste0(
      sym_condition,
      "_E=", environment,
      "_IM=", interaction_multiplier
    ),
    environment = factor(
      environment,
      levels = c("x2", "x4", "x8", "x16")
    )
  ) %>%
  mutate(
    treatment = as.factor(
      treatment
    )
  )
```

```{r}
# Pivot task data long:
# 1) Select only relevant columns
host_over_time_parent_task_data <- time_data %>%
  select(
    SEED,
    treatment,
    sym_condition,
    interaction_multiplier,
    environment,
    update,
    CurUpdate_AND_NOT_in_host_parent_org_counts,
    CurUpdate_AND_in_host_parent_org_counts,
    CurUpdate_EQU_in_host_parent_org_counts,
    CurUpdate_NAND_in_host_parent_org_counts,
    CurUpdate_NOR_in_host_parent_org_counts,
    CurUpdate_NOT_in_host_parent_org_counts,
    CurUpdate_OR_NOT_in_host_parent_org_counts,
    CurUpdate_OR_in_host_parent_org_counts,
    CurUpdate_XOR_in_host_parent_org_counts
  ) %>%
  pivot_longer(
    c(
      CurUpdate_AND_NOT_in_host_parent_org_counts,
      CurUpdate_AND_in_host_parent_org_counts,
      CurUpdate_EQU_in_host_parent_org_counts,
      CurUpdate_NAND_in_host_parent_org_counts,
      CurUpdate_NOR_in_host_parent_org_counts,
      CurUpdate_NOT_in_host_parent_org_counts,
      CurUpdate_OR_NOT_in_host_parent_org_counts,
      CurUpdate_OR_in_host_parent_org_counts,
      CurUpdate_XOR_in_host_parent_org_counts
    ),
    names_to = "task",
    values_to = "task_in_parent_count"
  ) %>%
  mutate(
    task = str_replace(task, "CurUpdate_", "")
  ) %>%
  mutate(
    task = str_replace(task, "_in_host_parent_org_counts", "")
  ) %>%
  mutate(
    task_completed_by_parent = task_in_parent_count >= threshold,
    task = factor(
      task,
      levels = tasks,
      labels = tasks
    )
  ) %>%
  mutate(
    task_complexity = case_when(
      task == "NOT" ~ 1,
      task == "NAND" ~ 1,
      task == "OR_NOT" ~ 2,
      task == "AND" ~ 2,
      task == "OR" ~ 3,
      task == "AND_NOT" ~ 3,
      task == "NOR" ~ 4,
      task == "XOR" ~ 4,
      task == "EQU" ~ 5,
      .default = -1
    )
  )

host_parent_tasks_all_time_by_task <- host_over_time_parent_task_data %>%
  group_by(
    SEED,
    treatment,
    sym_condition,
    interaction_multiplier,
    environment,
    task
  ) %>%
  summarize(
    n = n(),
    parent_task_completed_ever = any(task_completed_by_parent)
  ) %>%
  ungroup() %>%
  group_by(
    treatment,
    sym_condition,
    interaction_multiplier,
    environment,
    task
  ) %>%
  summarize(
    n = n(),
    parent_tasks_completed_ever = sum(parent_task_completed_ever == TRUE)
  )
```

Additionally, summarize by task complexity level.

```{r}
host_max_task_complexity_data <- host_over_time_parent_task_data %>%
  group_by(
    SEED,
    treatment,
    sym_condition,
    interaction_multiplier,
    environment,
    task,
    task_complexity
  ) %>%
  summarize(
    n = n(),
    parent_task_completed_ever = any(task_completed_by_parent)
  ) %>%
  ungroup() %>%
  # This filter works fine because all replicates start with a task
  filter(
    parent_task_completed_ever
  ) %>%
  group_by(
    SEED,
    treatment,
    sym_condition,
    interaction_multiplier,
    environment
  ) %>%
  summarize(
    max_task_complexity = max(task_complexity)
  ) %>%
  mutate(
    sym_condition = factor(
      sym_condition,
      levels = c(
        "no-syms",
        "parasite",
        "mutualist"
      )
    )
  )
```

## Calculate baseline for constraint

Use all no-symbiont replicates.

Calculate a bootstrapped 99% confidence interval for mean function complexity achieved
in the absence of symbionts.

```{r}
no_sym_max_task_complexity_ci <- host_max_task_complexity_data %>%
  ungroup() %>%
  filter(sym_condition == "no-syms" & interaction_multiplier == "4") %>%
  group_by(
    environment
  ) %>%
  summarize(
    ci = list(
      mean_cl_boot(
        max_task_complexity,
        conf.int = 0.95,
        B = 10000,
      ) %>%
      rename(
        no_sym_mean_max_task_complexity = y,
        no_sym_ci99_lwr_max_task_complexity = ymin,
        no_sym_ci99_upr_max_task_complexity = ymax
      )
    )
  ) %>%
  unnest(cols = c(ci))
```

```{r}
# library(boot)
# nonpara_temp <- host_max_task_complexity_data %>%
#   ungroup() %>%
#   filter(
#     sym_condition == "no-syms" & interaction_multiplier == "4" & environment == "x8"
#   ) %>%
#   group_by(
#     environment
#   )
# boot_out <- boot(
#   nonpara_temp$max_task_complexity,

# )
# boot.ci()
```

Next, categorize each replicate of the mutualism condition (across synergy levels) as constraining, promoting, or neither.

```{r}
mutualism_effect_on_hosts <- host_max_task_complexity_data %>%
  ungroup() %>%
  filter(sym_condition == "mutualist" & interaction_multiplier == "4") %>%
  left_join(no_sym_max_task_complexity_ci, by = "environment") %>%
  mutate(
    constrains = max_task_complexity < no_sym_ci99_lwr_max_task_complexity,
    promotes = max_task_complexity > no_sym_ci99_upr_max_task_complexity
  ) %>%
  mutate(
    mutualist_effect = case_when(
      constrains ~ "constraint",
      promotes ~ "promotes",
      .default = "none"
    )
  ) %>%
  group_by(
    environment,
    sym_condition
  ) %>%
  summarize(
    mean_max_task_complexity = mean(max_task_complexity),
    constrains_count = sum(constrains),
    promotes_count = sum(promotes),
    no_effect_count = sum(!constrains & !promotes)
  )
```

## Plot effects across synergy levels

```{r}
environment_constraint_plt <- mutualism_effect_on_hosts %>%
  ggplot(
    aes(
      x = environment,
      color = environment,
      fill = environment,
      y = constrains_count
    )
  ) +
  geom_col(
    position = "dodge",
    alpha = 0.5
  ) +
  geom_text(
    aes(label = constrains_count),
    position = position_dodge(width = 1),
    vjust = -0.5,
    size = 2
  ) +
  scale_color_dark() +
  scale_fill_dark() +
  scale_x_discrete() +
  scale_y_continuous(
    limits = c(0, 51)
  ) +
  theme(
    legend.position = "none"
  )

ggsave(
  file = paste(plot_dir, "environment_constraint_plt.pdf", sep = "/"),
  plot = environment_constraint_plt,
  width = 4,
  height = 4
)
```