# 2026-02-24 - Health mode synergy

## Dependencies and setup

```{r}
library(tidyverse)
library(cowplot)
library(RColorBrewer)
library(khroma)
library(rstatix)
library(knitr)
library(kableExtra)
library(viridis)
# library(webshot2)
```

```{r}
# Build relative path from R's working directory to what is actually the working
# directory for this experiment.
experiment_slug <- "2026-02-24-health-synergy"
working_directory <- paste(
  "experiments",
  experiment_slug,
  "analysis",
  sep = "/"
)
```

```{r}
# Use the cowplot theme as our default for ggplots
theme_set(theme_cowplot())
# Create directory to save plots
plot_dir <- paste(
  working_directory,
  "plots",
  sep = "/"
)
dir.create(
  plot_dir,
  showWarnings = FALSE
)
```

### Load and preprocess data

```{r}
tasks <- c("NOT", "NAND", "OR_NOT", "AND", "OR", "AND_NOT", "NOR", "XOR", "EQU")
max_pop_size <- 100 * 100
threshold <- 0.025 * max_pop_size
```

Load summary data.

```{r}
summary_data_path <- paste(
  working_directory,
  "data",
  "summary.csv",
  sep = "/"
)
summary_data <- read_csv(summary_data_path)
# Do any universal preprocessing
summary_data <- summary_data %>%
  mutate(
    SEED = as.factor(SEED),
    START_MOI = as.factor(START_MOI),
    VERTICAL_TRANSMISSION = as.factor(VERTICAL_TRANSMISSION),
    HEALTH_TYPE = as.factor(HEALTH_TYPE),
    HEALTH_INTERACTION_CHANCE = as.factor(HEALTH_INTERACTION_CHANCE),
    MUTUALIST_CYCLE_GAIN_PROP = as.factor(MUTUALIST_CYCLE_GAIN_PROP),
    PARASITE_CYCLE_LOSS_PROP = as.factor(PARASITE_CYCLE_LOSS_PROP),
    PARASITE_BASE_CYCLE_PROP = as.factor(PARASITE_BASE_CYCLE_PROP),
    MUTUALIST_CYCLE_DONATE_MULTIPLIER = as.factor(MUTUALIST_CYCLE_DONATE_MULTIPLIER), # nolint: line_length_linter.
    PARASITE_CYCLE_STEAL_MULTIPLIER = as.factor(PARASITE_CYCLE_STEAL_MULTIPLIER), # nolint: line_length_linter.
    HOST_ONLY_FIRST_TASK_CREDIT = as.factor(HOST_ONLY_FIRST_TASK_CREDIT),
    SYM_ONLY_FIRST_TASK_CREDIT = as.factor(SYM_ONLY_FIRST_TASK_CREDIT)
  ) %>%
  mutate(
    task_credit = paste0(
      "H", HOST_ONLY_FIRST_TASK_CREDIT,
      "S", SYM_ONLY_FIRST_TASK_CREDIT
    ),
    sym_condition = case_when(
      START_MOI == "0" ~ "no-syms",
      START_MOI == "1" & HEALTH_TYPE == "parasite" ~ "parasite",
      START_MOI == "1" & HEALTH_TYPE == "mutualist" ~ "mutualist",
      .default = "unknown" # precaution against unexpected parameterizations
    ),
    interaction_multiplier = as.factor(MUTUALIST_CYCLE_DONATE_MULTIPLIER),
    extinct_syms = OrgCounts_hosted_sym_count == 0
  ) %>%
  mutate(
    treatment = paste0(
      sym_condition,
      "_IM=", interaction_multiplier
    )
  ) %>%
  mutate(
    treatment = as.factor(
      treatment
    ),
    sym_con_extinction = as.factor(
      paste0(sym_condition, "_extinct=", extinct_syms)
    )
  )
```

```{r}
seeds_with_extinct_syms <- filter(
  summary_data,
  extinct_syms == TRUE
)$SEED
# Kind of gross way to "unfactor" the filtered list...
# Probably a better way to do it!
seeds_with_extinct_syms <- as.numeric(paste(seeds_with_extinct_syms))
```

Load time series data.

```{r}
time_data_path <- paste(
  working_directory,
  "data",
  "time_series.csv",
  sep = "/"
)
time_data <- read_csv(time_data_path)

# Do any universal preprocessing
time_data <- time_data %>%
  mutate(
    SEED = as.factor(SEED),
    START_MOI = as.factor(START_MOI),
    VERTICAL_TRANSMISSION = as.factor(VERTICAL_TRANSMISSION),
    HEALTH_TYPE = as.factor(HEALTH_TYPE),
    MUTUALIST_CYCLE_DONATE_MULTIPLIER = as.factor(MUTUALIST_CYCLE_DONATE_MULTIPLIER) # nolint
  ) %>%
  mutate(
    sym_condition = case_when(
      START_MOI == "0" ~ "no-syms",
      START_MOI == "1" & HEALTH_TYPE == "parasite" ~ "parasite",
      START_MOI == "1" & HEALTH_TYPE == "mutualist" ~ "mutualist",
      .default = "unknown" # precaution against unexpected parameterizations
    ),
    interaction_multiplier = as.factor(MUTUALIST_CYCLE_DONATE_MULTIPLIER),
    extinct_syms = SEED %in% seeds_with_extinct_syms
  ) %>%
  mutate(
    treatment = paste0(
      sym_condition,
      "_IM=", interaction_multiplier
    ),
    sym_con_extinction = as.factor(
      paste0(sym_condition, "_extinct=", extinct_syms)
    )
  ) %>%
  mutate(
    treatment = as.factor(
      treatment
    )
  )
```


## Which runs had hosts and/or symbiont populations that went extinct?

```{r}
sym_min_pop_sizes <- time_data %>%
  group_by(SEED) %>%
  slice_min(
    OrgCounts_hosted_sym_count,
    n = 1
  ) %>%
  slice_min(
    update,
    n = 1
  )

reps_with_extinct_syms <- sym_min_pop_sizes %>%
  group_by(
    treatment
  ) %>%
  summarize(
    reps_with_extinct_syms = sum(OrgCounts_hosted_sym_count == 0)
  )

host_min_pop_sizes <- time_data %>%
  group_by(SEED) %>%
  slice_min(
    OrgCounts_host_count,
    n = 1
  ) %>%
  slice_min(
    update,
    n = 1
  )

reps_with_extinct_hosts <- host_min_pop_sizes %>%
  group_by(
    treatment
  ) %>%
  summarize(
    reps_with_extinct_hosts = sum(OrgCounts_host_count == 0)
  )

extinctions <- full_join(
  reps_with_extinct_syms,
  reps_with_extinct_hosts
)
extinctions <- extinctions %>%
  arrange(
    desc(reps_with_extinct_hosts), desc(reps_with_extinct_syms)
  )
extinction_table <- kable(extinctions) %>%
  kable_styling(latex_options = "striped")
save_kable(
  extinction_table,
  paste(
    plot_dir,
    "extinctions.pdf",
    sep = "/"
  )
)
```

## Task performance all time

### Hosts

Quantify whether tasks have been performed over course of entire run.

```{r}
# Pivot task data long:
# 1) Select only relevant columns
host_over_time_parent_task_data <- time_data %>%
  select(
    SEED,
    treatment,
    sym_condition,
    sym_con_extinction,
    interaction_multiplier,
    update,
    CurUpdate_AND_NOT_in_host_parent_org_counts,
    CurUpdate_AND_in_host_parent_org_counts,
    CurUpdate_EQU_in_host_parent_org_counts,
    CurUpdate_NAND_in_host_parent_org_counts,
    CurUpdate_NOR_in_host_parent_org_counts,
    CurUpdate_NOT_in_host_parent_org_counts,
    CurUpdate_OR_NOT_in_host_parent_org_counts,
    CurUpdate_OR_in_host_parent_org_counts,
    CurUpdate_XOR_in_host_parent_org_counts
  ) %>%
  pivot_longer(
    c(
      CurUpdate_AND_NOT_in_host_parent_org_counts,
      CurUpdate_AND_in_host_parent_org_counts,
      CurUpdate_EQU_in_host_parent_org_counts,
      CurUpdate_NAND_in_host_parent_org_counts,
      CurUpdate_NOR_in_host_parent_org_counts,
      CurUpdate_NOT_in_host_parent_org_counts,
      CurUpdate_OR_NOT_in_host_parent_org_counts,
      CurUpdate_OR_in_host_parent_org_counts,
      CurUpdate_XOR_in_host_parent_org_counts
    ),
    names_to = "task",
    values_to = "task_in_parent_count"
  ) %>%
  mutate(
    task = str_replace(task, "CurUpdate_", "")
  ) %>%
  mutate(
    task = str_replace(task, "_in_host_parent_org_counts", "")
  ) %>%
  mutate(
    task_completed_by_parent = task_in_parent_count >= threshold,
    task = factor(
      task,
      levels = tasks,
      labels = tasks
    )
  ) %>%
  mutate(
    task_complexity = case_when(
      task == "NOT" ~ 1,
      task == "NAND" ~ 1,
      task == "OR_NOT" ~ 2,
      task == "AND" ~ 2,
      task == "OR" ~ 3,
      task == "AND_NOT" ~ 3,
      task == "NOR" ~ 4,
      task == "XOR" ~ 4,
      task == "EQU" ~ 5,
      .default = -1
    )
  )

host_parent_tasks_all_time_by_task <- host_over_time_parent_task_data %>%
  group_by(
    SEED,
    treatment,
    sym_condition,
    sym_con_extinction,
    interaction_multiplier,
    task
  ) %>%
  summarize(
    n = n(),
    parent_task_completed_ever = any(task_completed_by_parent)
  ) %>%
  ungroup() %>%
  group_by(
    treatment,
    sym_condition,
    interaction_multiplier,
    task
  ) %>%
  summarize(
    n = n(),
    parent_tasks_completed_ever = sum(parent_task_completed_ever == TRUE)
  )

```

Additionally, summarize by task complexity level.

```{r}
host_max_task_complexity_data <- host_over_time_parent_task_data %>%
  group_by(
    SEED,
    treatment,
    sym_condition,
    sym_con_extinction,
    interaction_multiplier,
    task,
    task_complexity
  ) %>%
  summarize(
    n = n(),
    parent_task_completed_ever = any(task_completed_by_parent)
  ) %>%
  ungroup() %>%
  # This filter works fine because all replicates start with a task
  filter(
    parent_task_completed_ever
  ) %>%
  group_by(
    SEED,
    treatment,
    sym_condition,
    sym_con_extinction,
    interaction_multiplier
  ) %>%
  summarize(
    max_task_complexity = max(task_complexity)
  ) %>%
  mutate(
    sym_condition = factor(
      sym_condition,
      levels = c(
        "no-syms",
        "parasite",
        "mutualist"
      )
    )
  )


```

Task complexity over time

```{r}
host_max_task_complexity_over_time_data <- host_over_time_parent_task_data %>%
  # This filter works fine because all replicates start with a task
  filter(
    task_completed_by_parent
  ) %>%
  group_by(
    SEED,
    treatment,
    sym_condition,
    interaction_multiplier,
    update
  ) %>%
  summarize(
    max_task_complexity = max(task_complexity)
  ) %>%
  mutate(
    sym_condition = factor(
      sym_condition,
      levels = c(
        "no-syms",
        "parasite",
        "mutualist"
      )
    )
  )
```

```{r}
host_parent_tasks_all_time_by_task <- host_parent_tasks_all_time_by_task %>%
  mutate(
    sym_condition = factor(
      sym_condition,
      levels = c(
        "no-syms",
        "parasite",
        "mutualist"
      )
    )
  )
host_parent_tasks_all_time_by_task_plt <- host_parent_tasks_all_time_by_task %>%
  ggplot(
    aes(
      x = task,
      color = sym_condition,
      fill = sym_condition,
      y = parent_tasks_completed_ever
    )
  ) +
  geom_col(
    position = "dodge",
    alpha = 0.5
  ) +
  geom_text(
    aes(label = parent_tasks_completed_ever),
    position = position_dodge(width = 1),
    vjust = -0.5,
    size = 2
  ) +
  scale_color_bright() +
  scale_fill_bright() +
  scale_x_discrete() +
  scale_y_continuous(
    limits = c(0, 51)
  ) +
  facet_wrap(
    ~interaction_multiplier,
    axes = "all_x",
    labeller = label_both
  ) +
  theme(
    legend.position = "bottom",
    axis.text.x = element_text(
      angle = 90,
      hjust = 1
    )
  )
ggsave(
  file = paste(plot_dir, "host_parent_tasks_all_time_by_task_plt.pdf", sep = "/"),
  plot = host_parent_tasks_all_time_by_task_plt,
  width = 15,
  height = 15
)
```

Total different tasks (all time)

```{r}
host_total_different_parent_tasks_all_time <- host_over_time_parent_task_data %>%
  group_by(
    SEED,
    treatment,
    sym_condition,
    interaction_multiplier,
    task
  ) %>%
  # Summarize across updates whether task was ever completed by a parent.
  summarize(
    n = n(),
    parent_task_completed_ever = any(task_completed_by_parent)
  ) %>%
  ungroup() %>%
  group_by(
    SEED,
    treatment,
    sym_condition,
    interaction_multiplier,
  ) %>%
  # Summarize across tasks (by seed)
  summarize(
    n = n(),
    num_parent_tasks_completed_ever = sum(parent_task_completed_ever == TRUE)
  )
```

Average number of different tasks evolved?

```{r}
host_avg_tasks_table <- host_total_different_parent_tasks_all_time %>%
  group_by(treatment) %>%
  summarize(avg_diff_num_host_parent_tasks = mean(num_parent_tasks_completed_ever)) %>%
  arrange(desc(avg_diff_num_host_parent_tasks)) %>%
  kable() %>%
  kable_styling(latex_options = "striped")
save_kable(
  host_avg_tasks_table,
  paste(
    plot_dir,
    "host_avg_diff_num_host_parent_tasks.pdf",
    sep = "/"
  )
)
```

Maximum task complexity evolved

```{r}
host_max_task_complexity_plt <- host_max_task_complexity_data %>%
  ggplot(
    aes(x = sym_condition, fill = sym_condition, y = max_task_complexity)
  ) +
  geom_point(
    mapping=aes(color = sym_condition),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_x_discrete() +
  facet_wrap(~interaction_multiplier) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(
      angle = 60,
      hjust = 1
    )
  )
ggsave(
  file = paste(plot_dir, "host_max_task_complexity_plt.pdf", sep = "/"),
  plot = host_max_task_complexity_plt,
  width = 10,
  height = 10
)
```

Task complexity over time

```{r}
host_max_task_complexity_over_time_plt <- host_max_task_complexity_over_time_data %>%
  filter(update %% 500 == 0) %>%
  ggplot(
    aes(
      x = update,
      y = max_task_complexity,
      color = sym_condition,
      fill = sym_condition
    )
  ) +
  stat_summary(fun = "mean", geom = "line") +
  stat_summary(
    fun.data = "mean_cl_boot",
    fun.args = list(conf.int = 0.95),
    geom = "ribbon",
    alpha = 0.2,
    linetype = 0
  ) +
  scale_color_bright() +
  scale_fill_bright() +
  facet_wrap(
    ~interaction_multiplier,
    axes = "all_x",
    labeller = label_both
  ) +
  theme(
    legend.position = "bottom"
  )

ggsave(
  file = paste(
    plot_dir,
    "host_max_task_complexity_over_time_plt.pdf",
    sep = "/"
  ),
  plot = host_max_task_complexity_over_time_plt,
  width = 15,
  height = 15
)
```

### Symbionts

Quantify whether tasks have been performed over course of entire run.

```{r}
# Pivot task data long:
# 1) Select only relevant columns
over_time_sym_parent_task_data <- time_data %>%
  select(
    SEED,
    treatment,
    sym_condition,
    interaction_multiplier,
    update,
    CurUpdate_AND_NOT_in_sym_parent_org_counts,
    CurUpdate_AND_in_sym_parent_org_counts,
    CurUpdate_EQU_in_sym_parent_org_counts,
    CurUpdate_NAND_in_sym_parent_org_counts,
    CurUpdate_NOR_in_sym_parent_org_counts,
    CurUpdate_NOT_in_sym_parent_org_counts,
    CurUpdate_OR_NOT_in_sym_parent_org_counts,
    CurUpdate_OR_in_sym_parent_org_counts,
    CurUpdate_XOR_in_sym_parent_org_counts
  ) %>%
  pivot_longer(
    c(
      CurUpdate_AND_NOT_in_sym_parent_org_counts,
      CurUpdate_AND_in_sym_parent_org_counts,
      CurUpdate_EQU_in_sym_parent_org_counts,
      CurUpdate_NAND_in_sym_parent_org_counts,
      CurUpdate_NOR_in_sym_parent_org_counts,
      CurUpdate_NOT_in_sym_parent_org_counts,
      CurUpdate_OR_NOT_in_sym_parent_org_counts,
      CurUpdate_OR_in_sym_parent_org_counts,
      CurUpdate_XOR_in_sym_parent_org_counts
    ),
    names_to = "task",
    values_to = "task_in_parent_count"
  ) %>%
  mutate(
    task = str_replace(task, "CurUpdate_", "")
  ) %>%
  mutate(
    task = str_replace(task, "_in_sym_parent_org_counts", "")
  ) %>%
  mutate(
    task_completed_by_parent = task_in_parent_count >= threshold,
    task = factor(
      task,
      levels = tasks,
      labels = tasks
    )
  )

sym_parent_tasks_all_time_by_task <- over_time_sym_parent_task_data %>%
  group_by(
    SEED,
    treatment,
    sym_condition,
    interaction_multiplier,
    task
  ) %>%
  summarize(
    n = n(),
    parent_task_completed_ever = any(task_completed_by_parent)
  ) %>%
  ungroup() %>%
  group_by(
    treatment,
    sym_condition,
    interaction_multiplier,
    task
  ) %>%
  summarize(
    n = n(),
    parent_tasks_completed_ever = sum(parent_task_completed_ever == TRUE)
  )

# parent_tasks_all_time_total <- parent_tasks_all_time_by_task %>%

```

```{r}
sym_parent_tasks_all_time_by_task <- sym_parent_tasks_all_time_by_task %>%
  mutate(
    sym_condition = factor(
      sym_condition,
      levels = c(
        "no-syms",
        "parasite",
        "mutualist"
      )
    )
  )
sym_parent_tasks_all_time_by_task_plt <- sym_parent_tasks_all_time_by_task %>%
  ggplot(
    aes(
      x = task,
      color = sym_condition,
      fill = sym_condition,
      y = parent_tasks_completed_ever
    )
  ) +
  geom_col(
    position = "dodge",
    alpha = 0.5
  ) +
  geom_text(
    aes(label = parent_tasks_completed_ever),
    position = position_dodge(width = 1),
    vjust = -0.5,
    size = 2
  ) +
  scale_color_bright() +
  scale_fill_bright() +
  scale_x_discrete() +
  scale_y_continuous(
    limits = c(0, 51)
  ) +
  facet_wrap(
    ~interaction_multiplier,
    axes = "all_x",
    labeller = label_both
  ) +
  theme(
    legend.position = "bottom",
    axis.text.x = element_text(
      angle = 90,
      hjust = 1
    )
  )
ggsave(
  file = paste(plot_dir, "sym_parent_tasks_all_time_by_task_plt.pdf", sep = "/"),
  plot = sym_parent_tasks_all_time_by_task_plt,
  width = 15,
  height = 15
)
```

## Phenotypic diversity

### Host diversity

#### Host number of task sets

```{r}
host_parent_num_task_sets_plt <- time_data %>%
  filter(update %% 500 == 0) %>%
  ggplot(
    aes(
      x = update,
      y = CurUpdate_host_parent_num_task_sets,
      color = sym_condition,
      fill = sym_condition
    )
  ) +
  stat_summary(fun = "mean", geom = "line") +
  stat_summary(
    fun.data = "mean_cl_boot",
    fun.args = list(conf.int = 0.95),
    geom = "ribbon",
    alpha = 0.2,
    linetype = 0
  ) +
  scale_color_bright() +
  scale_fill_bright() +
  facet_wrap(
    ~interaction_multiplier,
    axes = "all_x",
    labeller = label_both
  ) +
  theme(
    legend.position = "bottom"
  )
ggsave(
  file = paste(plot_dir, "host_parent_num_task_sets_plt.pdf", sep = "/"),
  plot = host_parent_num_task_sets_plt,
  width = 15,
  height = 15
)
```

#### Host entropy of task sets

```{r}
host_parent_entropy_task_sets_plt <- time_data %>%
  filter(update %% 500 == 0) %>%
  ggplot(
    aes(
      x = update,
      y = CurUpdate_host_parent_entropy_task_sets,
      color = sym_condition,
      fill = sym_condition
    )
  ) +
  stat_summary(fun = "mean", geom = "line") +
  stat_summary(
    fun.data = "mean_cl_boot",
    fun.args = list(conf.int = 0.95),
    geom = "ribbon",
    alpha = 0.2,
    linetype = 0
  ) +
  scale_color_bright() +
  scale_fill_bright() +
  facet_wrap(
    ~interaction_multiplier,
    axes = "all_x",
    labeller = label_both
  ) +
  theme(
    legend.position = "bottom"
  )
ggsave(
  file = paste(plot_dir, "host_parent_entropy_task_sets_plt.pdf", sep = "/"),
  plot = host_parent_entropy_task_sets_plt,
  width = 15,
  height = 15
)
```

### Symbiont diversity

#### Symbiont number of task sets

```{r}
symbiont_parent_num_task_sets_plt <- time_data %>%
  filter(update %% 500 == 0) %>%
  ggplot(
    aes(
      x = update,
      y = CurUpdate_sym_parent_num_task_sets,
      color = sym_condition,
      fill = sym_condition
    )
  ) +
  stat_summary(fun = "mean", geom = "line") +
  stat_summary(
    fun.data = "mean_cl_boot",
    fun.args = list(conf.int = 0.95),
    geom = "ribbon",
    alpha = 0.2,
    linetype = 0
  ) +
  scale_color_bright() +
  scale_fill_bright() +
  facet_wrap(
    ~interaction_multiplier,
    axes = "all_x",
    labeller = label_both
  ) +
  theme(
    legend.position = "bottom"
  )
ggsave(
  file = paste(plot_dir, "symbiont_parent_num_task_sets_plt.pdf", sep = "/"),
  plot = symbiont_parent_num_task_sets_plt,
  width = 15,
  height = 15
)
```

#### Symbiont entropy of task sets

```{r}
symbiont_parent_entropy_task_sets_plt <- time_data %>%
  filter(update %% 500 == 0) %>%
  ggplot(
    aes(
      x = update,
      y = CurUpdate_sym_parent_entropy_task_sets,
      color = sym_condition,
      fill = sym_condition
    )
  ) +
  stat_summary(fun = "mean", geom = "line") +
  stat_summary(
    fun.data = "mean_cl_boot",
    fun.args = list(conf.int = 0.95),
    geom = "ribbon",
    alpha = 0.2,
    linetype = 0
  ) +
  scale_color_bright() +
  scale_fill_bright() +
  facet_wrap(
    ~interaction_multiplier,
    axes = "all_x",
    labeller = label_both
  ) +
  theme(
    legend.position = "bottom"
  )
ggsave(
  file = paste(plot_dir, "symbiont_parent_entropy_task_sets_plt.pdf", sep = "/"),
  plot = symbiont_parent_entropy_task_sets_plt,
  width = 15,
  height = 15
)
```

## Transmission

### Attempts horizontal transmission

```{r}
TransmissionRates_attempts_horiztrans_plt <- time_data %>%
  filter(update %% 500 == 0) %>%
  ggplot(
    aes(
      x = update,
      y = TransmissionRates_attempts_horiztrans,
      color = sym_condition,
      fill = sym_condition
    )
  ) +
  stat_summary(fun = "mean", geom = "line") +
  stat_summary(
    fun.data = "mean_cl_boot",
    fun.args = list(conf.int = 0.95),
    geom = "ribbon",
    alpha = 0.2,
    linetype = 0
  ) +
  scale_color_bright() +
  scale_fill_bright() +
  facet_wrap(
    ~interaction_multiplier,
    axes = "all_x",
    labeller = label_both
  ) +
  theme(
    legend.position = "bottom"
  )
ggsave(
  file = paste(plot_dir, "TransmissionRates_attempts_horiztrans_plt.pdf", sep = "/"),
  plot = TransmissionRates_attempts_horiztrans_plt,
  width = 15,
  height = 15
)
```

```{r}

TransmissionRates_attempts_horiztrans_ALL_REPS_plt <- time_data %>%
  # group_by(
  #   sym_condition,
  #   interaction_multiplier
  # ) %>%
  # arrange(SEED) %>%
  mutate(rep_id = as.numeric(SEED) %% 50) %>%
  # ungroup() %>%
  filter(update %% 500 == 0 & sym_condition == "mutualist") %>%
  ggplot(
    aes(
      x = update,
      y = TransmissionRates_attempts_horiztrans,
    )
  ) +
  geom_line() +
  facet_grid(
    rep_id ~ interaction_multiplier,
    scales = "free_y"
  )
ggsave(
  file = paste(plot_dir, "TransmissionRates_attempts_horiztrans_ALL_REPS_plt.pdf", sep = "/"),
  plot = TransmissionRates_attempts_horiztrans_ALL_REPS_plt,
  width = 20,
  height = 40
)
```

### Successes horizontal transmission

```{r}
TransmissionRates_successes_horiztrans_plt <- time_data %>%
  filter(update %% 500 == 0) %>%
  ggplot(
    aes(
      x = update,
      y = TransmissionRates_successes_horiztrans,
      color = sym_condition,
      fill = sym_condition
    )
  ) +
  stat_summary(fun = "mean", geom = "line") +
  stat_summary(
    fun.data = "mean_cl_boot",
    fun.args = list(conf.int = 0.95),
    geom = "ribbon",
    alpha = 0.2,
    linetype = 0
  ) +
  scale_color_bright() +
  scale_fill_bright() +
  facet_wrap(
    ~interaction_multiplier,
    axes = "all_x",
    labeller = label_both
  ) +
  theme(
    legend.position = "bottom"
  )
ggsave(
  file = paste(plot_dir, "TransmissionRates_successes_horiztrans_plt.pdf", sep = "/"),
  plot = TransmissionRates_successes_horiztrans_plt,
  width = 15,
  height = 15
)
```

### Attempts vertical transmission

```{r}
TransmissionRates_attempts_verttrans_plt <- time_data %>%
  filter(update %% 500 == 0) %>%
  ggplot(
    aes(
      x = update,
      y = TransmissionRates_attempts_verttrans,
      color = sym_condition,
      fill = sym_condition
    )
  ) +
  stat_summary(fun = "mean", geom = "line") +
  stat_summary(
    fun.data = "mean_cl_boot",
    fun.args = list(conf.int = 0.95),
    geom = "ribbon",
    alpha = 0.2,
    linetype = 0
  ) +
  scale_color_bright() +
  scale_fill_bright() +
  facet_wrap(
    ~interaction_multiplier,
    axes = "all_x",
    labeller = label_both
  ) +
  theme(
    legend.position = "bottom"
  )
ggsave(
  file = paste(plot_dir, "TransmissionRates_attempts_verttrans_plt.pdf", sep = "/"),
  plot = TransmissionRates_attempts_verttrans_plt,
  width = 15,
  height = 15
)
```

### Successes vertical transmission

```{r}
TransmissionRates_successes_verttrans_plt <- time_data %>%
  filter(update %% 500 == 0) %>%
  ggplot(
    aes(
      x = update,
      y = TransmissionRates_successes_verttrans,
      color = sym_condition,
      fill = sym_condition
    )
  ) +
  stat_summary(fun = "mean", geom = "line") +
  stat_summary(
    fun.data = "mean_cl_boot",
    fun.args = list(conf.int = 0.95),
    geom = "ribbon",
    alpha = 0.2,
    linetype = 0
  ) +
  scale_color_bright() +
  scale_fill_bright() +
  facet_wrap(
    ~interaction_multiplier,
    axes = "all_x",
    labeller = label_both
  ) +
  theme(
    legend.position = "bottom"
  )
ggsave(
  file = paste(plot_dir, "TransmissionRates_successes_verttrans_plt.pdf", sep = "/"),
  plot = TransmissionRates_successes_verttrans_plt,
  width = 15,
  height = 15
)
```


Summarize transmission rates at the end of the experiment

```{r}
trans_rates <- time_data %>%
  filter(sym_condition == "mutualist" & update >= 190000) %>%
  group_by(
    SEED,
    interaction_multiplier,
    treatment
  ) %>%
  summarise(
    total_ht_attempts = sum(TransmissionRates_attempts_horiztrans),
    total_ht_successes = sum(TransmissionRates_successes_horiztrans),
    total_vt_successes = sum(TransmissionRates_successes_verttrans),
    ht_prop = sum(TransmissionRates_successes_horiztrans) / (sum(TransmissionRates_successes_horiztrans) + sum(TransmissionRates_successes_verttrans)),
    vt_prop = sum(TransmissionRates_successes_verttrans) / (sum(TransmissionRates_successes_horiztrans) + sum(TransmissionRates_successes_verttrans))
  ) %>%
  group_by(interaction_multiplier) %>%
  summarise(
    # mean_ht_attempts = mean(total_ht_attempts),
    num_reps_with_ht = sum(total_ht_successes > 100),
    mean_ht_successes = mean(total_ht_successes),
    mean_vt_successes = mean(total_vt_successes),
    mean_ht_prop = mean(ht_prop),
    mean_vt_prop = mean(vt_prop)
  )
  # group_by(environment) %>%
  # summarise(
  #   num_ht_attempts = sum(TransmissionRates_attempts_horiztrans > 0)
  # )

## Population size

Average / median population size

```{r}
pop_size_summary <- time_data %>%
  group_by(
    SEED,
    treatment
  ) %>%
  summarize(
    host_pop_size_avg = mean(OrgCounts_host_count),
    host_pop_size_median = median(OrgCounts_host_count),
    host_pop_size_variance = var(OrgCounts_host_count),
    sym_pop_size_avg = mean(OrgCounts_hosted_sym_count),
    sym_pop_size_median = median(OrgCounts_hosted_sym_count),
    sym_pop_size_variance = var(OrgCounts_hosted_sym_count)
  )
```

```{r}
host_pop_size_avg_plt <- pop_size_summary %>%
  ggplot(
    aes(x = treatment, fill = treatment, y = host_pop_size_avg)
  ) +
  geom_point(
    mapping=aes(color = treatment),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_x_discrete() +
  # facet_wrap(STRESS_FREQUENCY~stress_death_chance, labeller = label_both) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(
      angle = 60,
      hjust = 1
    )
  )
ggsave(
  file = paste(plot_dir, "host_pop_size_avg_plt.pdf", sep = "/"),
  plot = host_pop_size_avg_plt,
  width = 10,
  height = 10
)
```

```{r}
host_pop_size_variance_plt <- pop_size_summary %>%
  ggplot(
    aes(x = treatment, fill = treatment, y = host_pop_size_variance)
  ) +
  geom_point(
    mapping=aes(color = treatment),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_x_discrete() +
  # facet_wrap(STRESS_FREQUENCY~stress_death_chance, labeller = label_both) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(
      angle = 60,
      hjust = 1
    )
  )
ggsave(
  file = paste(plot_dir, "host_pop_size_variance_plt.pdf", sep = "/"),
  plot = host_pop_size_variance_plt,
  width = 10,
  height = 10
)
```

```{r}
sym_pop_size_avg_plt <- pop_size_summary %>%
  ggplot(
    aes(x = treatment, fill = treatment, y = sym_pop_size_avg)
  ) +
  geom_point(
    mapping = aes(color = treatment),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_x_discrete() +
  # facet_wrap(STRESS_FREQUENCY~stress_death_chance, labeller = label_both) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(
      angle = 60,
      hjust = 1
    )
  )
ggsave(
  file = paste(plot_dir, "sym_pop_size_avg_plt.pdf", sep = "/"),
  plot = sym_pop_size_avg_plt,
  width = 10,
  height = 10
)
```

## Generations elapsed

### Hosts

```{r}
host_avg_gens_elapsed_plt <- summary_data %>%
  ggplot(
    aes(x = treatment, fill = treatment, y = CurUpdate_host_mean_generations)
  ) +
  geom_point(
    mapping=aes(color = treatment),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_x_discrete() +
  theme(
    legend.position = "none",
    axis.text.x = element_text(
      angle = 60,
      hjust = 1
    )
  )
ggsave(
  file = paste(plot_dir, "host_avg_gens_elapsed_plt.pdf", sep = "/"),
  plot = host_avg_gens_elapsed_plt,
  width = 10,
  height = 10
)
```

### Symbionts

```{r}
sym_avg_gens_elapsed_plt <- summary_data %>%
  ggplot(
    aes(x = treatment, fill = treatment, y = CurUpdate_sym_mean_generations)
  ) +
  geom_point(
    mapping=aes(color = treatment),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_x_discrete() +
  theme(
    legend.position = "none",
    axis.text.x = element_text(
      angle = 60,
      hjust = 1
    )
  )
ggsave(
  file = paste(plot_dir, "sym_avg_gens_elapsed_plt.pdf", sep = "/"),
  plot = sym_avg_gens_elapsed_plt,
  width = 10,
  height = 10
)
```

## Individual tasks over time

## Task performance over time

```{r}
over_time_sym_parent_task_data <- over_time_sym_parent_task_data %>%
  mutate(org_type = "symbiont")

host_over_time_parent_task_data <- host_over_time_parent_task_data %>%
  mutate(org_type = "host")

over_time_parent_task_data <- full_join(
  over_time_sym_parent_task_data,
  host_over_time_parent_task_data
)
over_time_parent_task_data <- over_time_parent_task_data %>%
  mutate(
    org_type = as.factor(org_type)
  )
```

```{r}
treatments <- levels(over_time_parent_task_data$treatment)

for (trt in treatments) {
  print(trt)
  p <- over_time_parent_task_data %>%
    filter(treatment == trt) %>%
    ggplot(
      aes(
        x = update,
        y = task_in_parent_count,
        color = org_type,
        fill = org_type
      )
    ) +
    geom_line() +
    labs(
      title = trt
    ) +
    facet_grid(
      SEED ~ task,
      scales = "free_y",
      labeller = label_both
    )

  ggsave(
    file = paste(
      plot_dir,
      paste0("tasks_over_time_", trt, ".pdf"),
      sep = "/"
    ),
    plot = p,
    width = 20,
    height = 40
  )
}
```

## Manuscript figures

```{r}

host_max_task_complexity_data_ci <- host_max_task_complexity_data %>%
  group_by(
    treatment,
    sym_condition,
    interaction_multiplier
  ) %>%
  summarize(
    mean_max_task_complexity = mean(max_task_complexity),
    min_max_task_complexity = min(max_task_complexity),
    max_max_task_complexity = max(max_task_complexity),
    ci = list(
      mean_cl_boot(
        max_task_complexity,
        conf.int = 0.95,
        B = 10000,
      ) %>% rename(mean=y, lwr=ymin, upr=ymax)
    )
  ) %>%
  unnest(cols = c(ci)) %>%
  filter(
    interaction_multiplier == "4"
  ) %>%
  mutate(
    sym_condition = factor(
      sym_condition,
      levels = c("no-syms", "parasite", "mutualist"),
      labels = c("No symbionts", "Parasites", "Mutualists")
    )
  )

manuscript_host_task_complexity_plt <- host_max_task_complexity_data_ci %>%
  ggplot(
    aes(
      x = sym_condition,
      fill = sym_condition,
      y = mean
    )
  ) +
  geom_bar(stat = "identity", alpha = 0.7) +
  geom_linerange(
    aes(
      ymin = lwr,
      ymax = upr
    ),
    alpha = 0.9,
    size = 0.5
  ) +
  scale_x_discrete() +
  scale_fill_bright() +
  scale_y_continuous(
    limits = c(0, 5.1)
  ) +
  labs(
    x = "Symbiont type",
    y = "Mean function complexity"
  ) +
  # facet_wrap(~interaction_multiplier) +
  theme(
    legend.position = "none"
    # axis.text.x = element_text(
    #   angle = 45,
    #   hjust = 1
    # )
  )
ggsave(
  file = paste(plot_dir, "manuscript_host_task_complexity_plt.pdf", sep = "/"),
  plot = manuscript_host_task_complexity_plt,
  width = 4,
  height = 4
)
```

```{r}
manuscript_host_max_task_complexity_over_time_plt <- host_max_task_complexity_over_time_data %>%
  filter(
    # update %% 10000 == 0 &
    interaction_multiplier == "4"
  ) %>%
  mutate(
    sym_condition = factor(
      sym_condition,
      levels = c("no-syms", "parasite", "mutualist"),
      labels = c("No symbionts", "Parasites", "Mutualists")
    )
  ) %>%
  ggplot(
    aes(
      x = update,
      y = max_task_complexity,
      color = sym_condition,
      fill = sym_condition
    )
  ) +
  stat_summary(fun = "mean", geom = "line") +
  stat_summary(
    fun.data = "mean_cl_boot",
    fun.args = list(conf.int = 0.95),
    geom = "ribbon",
    alpha = 0.2,
    linetype = 0
  ) +
  scale_color_bright() +
  scale_fill_bright() +
  scale_y_continuous(
    limits = c(0, 5.1)
  ) +
  labs(
    x = "Updates",
    y = "Mean function complexity",
    color = "Symbiont type",
    fill = "Symbiont type"
  ) +
  theme(
    legend.position = "bottom"
  )

ggsave(
  file = paste(
    plot_dir,
    "manuscript_host_max_task_complexity_over_time_plt.pdf",
    sep = "/"
  ),
  plot = manuscript_host_max_task_complexity_over_time_plt,
  width = 6,
  height = 4
)
```

Combined figure

```{r}
plt_grid <- plot_grid(
  manuscript_host_max_task_complexity_over_time_plt +
    theme(
      legend.position = "bottom",
      legend.title = element_text(size = 10),
      legend.text = element_text(size = 8),
      axis.title.y = element_text(size = 10),
      axis.text.y = element_text(size = 8),
      axis.title.x = element_text(size = 10),
      axis.text.x = element_text(size = 8)
      # plot.margin = margin(t = 1, unit = "cm")
    ),
  manuscript_host_task_complexity_plt +
    theme(
      legend.position = "none",
      axis.title.y = element_blank(), #element_text(size = 10),
      axis.text.y = element_text(size = 8),
      axis.title.x = element_text(size = 10),
      axis.text.x = element_text(size = 8)
      # plot.margin = margin(t = 1, l = 0.5, unit = "cm")
    ),
  nrow = 1,
  ncol = 2,
  align = "h",
  # labels = c("A)", "B)"),
  rel_widths = c(0.7, 0.3)
)
# plt_grid
save_plot(
  filename = paste(
    plot_dir,
    "fig-diff-rewards-host-complexity.pdf",
    sep = "/"
  ),
  plot = plt_grid,
  base_width = 8,
  base_height = 4,
  dpi = 600
)
```