# 2026-02-26 - Health mode evolvable interaction values - differential reward values

## Dependencies and setup

```{r}
library(tidyverse)
library(cowplot)
library(RColorBrewer)
library(khroma)
library(rstatix)
library(knitr)
library(kableExtra)
library(viridis)
library(ggh4x)
# library(webshot2)
```

```{r}
# Build relative path from R's working directory to what is actually the working
# directory for this experiment.
experiment_slug <- "2026-02-26-health-evo-intval-diff-rewards"
working_directory <- paste(
  "experiments",
  experiment_slug,
  "analysis",
  sep = "/"
)
```

```{r}
# Use the cowplot theme as our default for ggplots
theme_set(theme_cowplot())
# Create directory to save plots
plot_dir <- paste(
  working_directory,
  "plots",
  sep = "/"
)
dir.create(
  plot_dir,
  showWarnings = FALSE
)
```

### Load and preprocess data

```{r}
tasks <- c("NOT", "NAND", "OR_NOT", "AND", "OR", "AND_NOT", "NOR", "XOR", "EQU")
max_pop_size <- 100 * 100
threshold <- 0.025 * max_pop_size
num_replicates <- 50
```

Load summary data.

```{r}
summary_data_path <- paste(
  working_directory,
  "data",
  "summary.csv",
  sep = "/"
)
summary_data <- read_csv(summary_data_path)
# Do any universal preprocessing
summary_data <- summary_data %>%
  mutate(
    SEED = as.factor(SEED),
    START_MOI = as.factor(START_MOI),
    VERTICAL_TRANSMISSION = as.factor(VERTICAL_TRANSMISSION),
    MUTUALIST_CYCLE_DONATE_MULTIPLIER = as.factor(MUTUALIST_CYCLE_DONATE_MULTIPLIER), # nolint: line_length_linter.
    PARASITE_CYCLE_STEAL_MULTIPLIER = as.factor(PARASITE_CYCLE_STEAL_MULTIPLIER), # nolint: line_length_linter.
    HEALTH_TYPE = as.factor(HEALTH_TYPE),
    PARASITE_CYCLE_LOSS_PROP = as.factor(PARASITE_CYCLE_LOSS_PROP),
    HEALTH_INTERACTION_CHANCE = as.factor(HEALTH_INTERACTION_CHANCE),
    PARASITE_BASE_CYCLE_PROP = as.factor(PARASITE_BASE_CYCLE_PROP)
  ) %>%
  mutate(
    treatment = paste0(
      "VT=", VERTICAL_TRANSMISSION
    )
  ) %>%
  mutate(
    treatment = as.factor(
      treatment
    )
  )
```

Load time series data.

```{r}
time_data_path <- paste(
  working_directory,
  "data",
  "time_series.csv",
  sep = "/"
)
time_data <- read_csv(time_data_path)

# Do any universal preprocessing
time_data <- time_data %>%
  mutate(
    SEED = as.factor(SEED),
    VERTICAL_TRANSMISSION = as.factor(VERTICAL_TRANSMISSION)
  ) %>%
  mutate(
    treatment = paste0(
      "VT=", VERTICAL_TRANSMISSION
    )
  ) %>%
  mutate(
    treatment = as.factor(
      treatment
    )
  )
```


## Which runs had hosts and/or symbiont populations that went extinct?

```{r}
sym_min_pop_sizes <- time_data %>%
  group_by(SEED) %>%
  slice_min(
    OrgCounts_hosted_sym_count,
    n = 1
  ) %>%
  slice_min(
    update,
    n = 1
  )

reps_with_extinct_syms <- sym_min_pop_sizes %>%
  group_by(
    treatment
  ) %>%
  summarize(
    reps_with_extinct_syms = sum(OrgCounts_hosted_sym_count == 0)
  )

host_min_pop_sizes <- time_data %>%
  group_by(SEED) %>%
  slice_min(
    OrgCounts_host_count,
    n = 1
  ) %>%
  slice_min(
    update,
    n = 1
  )

reps_with_extinct_hosts <- host_min_pop_sizes %>%
  group_by(
    treatment
  ) %>%
  summarize(
    reps_with_extinct_hosts = sum(OrgCounts_host_count == 0)
  )

extinctions <- full_join(
  reps_with_extinct_syms,
  reps_with_extinct_hosts
)
extinctions <- extinctions %>%
  arrange(
    desc(reps_with_extinct_hosts), desc(reps_with_extinct_syms)
  )
extinction_table <- kable(extinctions) %>%
  kable_styling(latex_options = "striped")
save_kable(
  extinction_table,
  paste(
    plot_dir,
    "extinctions.pdf",
    sep = "/"
  )
)
```

## Task performance all time

### Hosts


Quantify whether tasks have been performed over course of entire run.

```{r}
# Pivot task data long:
# 1) Select only relevant columns
host_over_time_parent_task_data <- time_data %>%
  select(
    SEED,
    treatment,
    VERTICAL_TRANSMISSION,
    update,
    CurUpdate_AND_NOT_in_host_parent_org_counts,
    CurUpdate_AND_in_host_parent_org_counts,
    CurUpdate_EQU_in_host_parent_org_counts,
    CurUpdate_NAND_in_host_parent_org_counts,
    CurUpdate_NOR_in_host_parent_org_counts,
    CurUpdate_NOT_in_host_parent_org_counts,
    CurUpdate_OR_NOT_in_host_parent_org_counts,
    CurUpdate_OR_in_host_parent_org_counts,
    CurUpdate_XOR_in_host_parent_org_counts
  ) %>%
  pivot_longer(
    c(
      CurUpdate_AND_NOT_in_host_parent_org_counts,
      CurUpdate_AND_in_host_parent_org_counts,
      CurUpdate_EQU_in_host_parent_org_counts,
      CurUpdate_NAND_in_host_parent_org_counts,
      CurUpdate_NOR_in_host_parent_org_counts,
      CurUpdate_NOT_in_host_parent_org_counts,
      CurUpdate_OR_NOT_in_host_parent_org_counts,
      CurUpdate_OR_in_host_parent_org_counts,
      CurUpdate_XOR_in_host_parent_org_counts
    ),
    names_to = "task",
    values_to = "task_in_parent_count"
  ) %>%
  mutate(
    task = str_replace(task, "CurUpdate_", "")
  ) %>%
  mutate(
    task = str_replace(task, "_in_host_parent_org_counts", "")
  ) %>%
  mutate(
    task_completed_by_parent = task_in_parent_count >= threshold,
    task = factor(
      task,
      levels = tasks,
      labels = tasks
    )
  ) %>%
  mutate(
    task_complexity = case_when(
      task == "NOT" ~ 1,
      task == "NAND" ~ 1,
      task == "OR_NOT" ~ 2,
      task == "AND" ~ 2,
      task == "OR" ~ 3,
      task == "AND_NOT" ~ 3,
      task == "NOR" ~ 4,
      task == "XOR" ~ 4,
      task == "EQU" ~ 5,
      .default = -1
    )
  )

host_parent_tasks_all_time_by_task <- host_over_time_parent_task_data %>%
  group_by(
    SEED,
    treatment,
    VERTICAL_TRANSMISSION,
    task
  ) %>%
  summarize(
    n = n(),
    parent_task_completed_ever = any(task_completed_by_parent)
  ) %>%
  ungroup() %>%
  group_by(
    treatment,
    VERTICAL_TRANSMISSION,
    task
  ) %>%
  summarize(
    n = n(),
    parent_tasks_completed_ever = sum(parent_task_completed_ever == TRUE)
  )

```

Additionally, summarize by task complexity level.

```{r}
host_max_task_complexity_data <- host_over_time_parent_task_data %>%
  group_by(
    SEED,
    treatment,
    VERTICAL_TRANSMISSION,
    task,
    task_complexity
  ) %>%
  summarize(
    n = n(),
    parent_task_completed_ever = any(task_completed_by_parent)
  ) %>%
  ungroup() %>%
  # This filter works fine because all replicates start with a task
  filter(
    parent_task_completed_ever
  ) %>%
  group_by(
    SEED,
    treatment,
    VERTICAL_TRANSMISSION
  ) %>%
  summarize(
    max_task_complexity = max(task_complexity)
  )


```

Task complexity over time

```{r}
host_max_task_complexity_over_time_data <- host_over_time_parent_task_data %>%
  # This filter works fine because all replicates start with a task
  filter(
    task_completed_by_parent
  ) %>%
  group_by(
    SEED,
    treatment,
    VERTICAL_TRANSMISSION,
    update
  ) %>%
  summarize(
    max_task_complexity = max(task_complexity)
  )
```

#### Host parent tasks all time (by task)

```{r}
host_parent_tasks_all_time_by_task_plt <- host_parent_tasks_all_time_by_task %>%
  ggplot(
    aes(
      x = task,
      color = task,
      fill = task,
      y = parent_tasks_completed_ever
    )
  ) +
  geom_col(
    position = "dodge",
    alpha = 0.5
  ) +
  geom_text(
    aes(label = parent_tasks_completed_ever),
    position = position_dodge(width = 1),
    vjust = -0.5,
    size = 2
  ) +
  scale_x_discrete() +
  scale_y_continuous(
    limits = c(0, 51)
  ) +
  facet_wrap(
    ~ VERTICAL_TRANSMISSION,
    axes = "all_x",
    labeller = label_both
  ) +
  theme(
    legend.position = "bottom",
    axis.text.x = element_text(
      angle = 90,
      hjust = 1
    )
  )
ggsave(
  file = paste(plot_dir, "host_parent_tasks_all_time_by_task_plt.pdf", sep = "/"),
  plot = host_parent_tasks_all_time_by_task_plt,
  width = 15,
  height = 15
)
```

#### Host task complexity evolved

```{r}
host_max_task_complexity_plt <- host_max_task_complexity_data %>%
  ggplot(
    aes(
      x = VERTICAL_TRANSMISSION,
      fill = VERTICAL_TRANSMISSION,
      y = max_task_complexity
    )
  ) +
  geom_point(
    mapping=aes(color = VERTICAL_TRANSMISSION),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_x_discrete() +
  # facet_grid(
  #   ~ task_credit,
  #   axes = "all_x"
  # ) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(
      angle = 60,
      hjust = 1
    )
  )
ggsave(
  file = paste(plot_dir, "host_max_task_complexity_plt.pdf", sep = "/"),
  plot = host_max_task_complexity_plt,
  width = 10,
  height = 10
)
```

Task complexity over time

```{r}
host_max_task_complexity_over_time_plt <- host_max_task_complexity_over_time_data %>%
  filter(update %% 10000 == 0) %>%
  ggplot(
    aes(
      x = update,
      y = max_task_complexity,
      color = VERTICAL_TRANSMISSION,
      fill = VERTICAL_TRANSMISSION
    )
  ) +
  stat_summary(fun = "mean", geom = "line") +
  stat_summary(
    fun.data = "mean_cl_boot",
    fun.args = list(conf.int = 0.95),
    geom = "ribbon",
    alpha = 0.2,
    linetype = 0
  ) +
  theme(
    legend.position = "bottom"
  )

ggsave(
  file = paste(
    plot_dir,
    "host_max_task_complexity_over_time_plt.pdf",
    sep = "/"
  ),
  plot = host_max_task_complexity_over_time_plt,
  width = 15,
  height = 15
)
```


### Symbionts

Quantify whether tasks have been performed over course of entire run.

```{r}
# Pivot task data long:
# 1) Select only relevant columns
over_time_sym_parent_task_data <- time_data %>%
  select(
    SEED,
    treatment,
    VERTICAL_TRANSMISSION,
    update,
    CurUpdate_AND_NOT_in_sym_parent_org_counts,
    CurUpdate_AND_in_sym_parent_org_counts,
    CurUpdate_EQU_in_sym_parent_org_counts,
    CurUpdate_NAND_in_sym_parent_org_counts,
    CurUpdate_NOR_in_sym_parent_org_counts,
    CurUpdate_NOT_in_sym_parent_org_counts,
    CurUpdate_OR_NOT_in_sym_parent_org_counts,
    CurUpdate_OR_in_sym_parent_org_counts,
    CurUpdate_XOR_in_sym_parent_org_counts
  ) %>%
  pivot_longer(
    c(
      CurUpdate_AND_NOT_in_sym_parent_org_counts,
      CurUpdate_AND_in_sym_parent_org_counts,
      CurUpdate_EQU_in_sym_parent_org_counts,
      CurUpdate_NAND_in_sym_parent_org_counts,
      CurUpdate_NOR_in_sym_parent_org_counts,
      CurUpdate_NOT_in_sym_parent_org_counts,
      CurUpdate_OR_NOT_in_sym_parent_org_counts,
      CurUpdate_OR_in_sym_parent_org_counts,
      CurUpdate_XOR_in_sym_parent_org_counts
    ),
    names_to = "task",
    values_to = "task_in_parent_count"
  ) %>%
  mutate(
    task = str_replace(task, "CurUpdate_", "")
  ) %>%
  mutate(
    task = str_replace(task, "_in_sym_parent_org_counts", "")
  ) %>%
  mutate(
    task_completed_by_parent = task_in_parent_count >= threshold,
    task = factor(
      task,
      levels = tasks,
      labels = tasks
    )
  )

sym_parent_tasks_all_time_by_task <- over_time_sym_parent_task_data %>%
  group_by(
    SEED,
    treatment,
    VERTICAL_TRANSMISSION,
    task
  ) %>%
  summarize(
    n = n(),
    parent_task_completed_ever = any(task_completed_by_parent)
  ) %>%
  ungroup() %>%
  group_by(
    treatment,
    VERTICAL_TRANSMISSION,
    task
  ) %>%
  summarize(
    n = n(),
    parent_tasks_completed_ever = sum(parent_task_completed_ever == TRUE)
  )

```

```{r}

sym_parent_tasks_all_time_by_task_plt <- sym_parent_tasks_all_time_by_task %>%
  ggplot(
    aes(
      x = task,
      color = task,
      fill = task,
      y = parent_tasks_completed_ever
    )
  ) +
  geom_col(
    position = "dodge",
    alpha = 0.5
  ) +
  geom_text(
    aes(label = parent_tasks_completed_ever),
    position = position_dodge(width = 1),
    vjust = -0.5,
    size = 2
  ) +
  scale_x_discrete() +
  scale_y_continuous(
    limits = c(0, 51)
  ) +
  facet_wrap(
    ~ VERTICAL_TRANSMISSION,
    axes = "all_x",
    labeller = label_both
  ) +
  theme(
    legend.position = "bottom",
    axis.text.x = element_text(
      angle = 90,
      hjust = 1
    )
  )
ggsave(
  file = paste(plot_dir, "sym_parent_tasks_all_time_by_task_plt.pdf", sep = "/"),
  plot = sym_parent_tasks_all_time_by_task_plt,
  width = 15,
  height = 15
)
```

## Population size

Average / median population size

```{r}
pop_size_summary <- time_data %>%
  group_by(
    SEED,
    treatment
  ) %>%
  summarize(
    host_pop_size_avg = mean(OrgCounts_host_count),
    host_pop_size_median = median(OrgCounts_host_count),
    host_pop_size_variance = var(OrgCounts_host_count),
    sym_pop_size_avg = mean(OrgCounts_hosted_sym_count),
    sym_pop_size_median = median(OrgCounts_hosted_sym_count),
    sym_pop_size_variance = var(OrgCounts_hosted_sym_count)
  )
```

```{r}
host_pop_size_avg_plt <- pop_size_summary %>%
  ggplot(
    aes(x = treatment, fill = treatment, y = host_pop_size_avg)
  ) +
  geom_point(
    mapping=aes(color = treatment),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_x_discrete() +
  # facet_wrap(STRESS_FREQUENCY~stress_death_chance, labeller = label_both) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(
      angle = 60,
      hjust = 1
    )
  )
ggsave(
  file = paste(plot_dir, "host_pop_size_avg_plt.pdf", sep = "/"),
  plot = host_pop_size_avg_plt,
  width = 10,
  height = 10
)
```

```{r}
host_pop_size_variance_plt <- pop_size_summary %>%
  ggplot(
    aes(x = treatment, fill = treatment, y = host_pop_size_variance)
  ) +
  geom_point(
    mapping=aes(color = treatment),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_x_discrete() +
  # facet_wrap(STRESS_FREQUENCY~stress_death_chance, labeller = label_both) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(
      angle = 60,
      hjust = 1
    )
  )
ggsave(
  file = paste(plot_dir, "host_pop_size_variance_plt.pdf", sep = "/"),
  plot = host_pop_size_variance_plt,
  width = 10,
  height = 10
)
```

## Task performance over time

```{r}
over_time_sym_parent_task_data <- over_time_sym_parent_task_data %>%
  mutate(org_type = "symbiont")

host_over_time_parent_task_data <- host_over_time_parent_task_data %>%
  mutate(org_type = "host")

over_time_parent_task_data <- full_join(
  over_time_sym_parent_task_data,
  host_over_time_parent_task_data
)
over_time_parent_task_data <- over_time_parent_task_data %>%
  mutate(
    org_type = as.factor(org_type)
  )
```

```{r}
treatments <- levels(over_time_parent_task_data$treatment)

for (trt in treatments) {
  print(trt)
  p <- over_time_parent_task_data %>%
    filter(treatment == trt) %>%
    ggplot(
      aes(
        x = update,
        y = task_in_parent_count,
        color = org_type,
        fill = org_type
      )
    ) +
    geom_line() +
    labs(
      title = trt
    ) +
    facet_grid(
      SEED ~ task,
      scales = "free_y",
      labeller = label_both
    )

  ggsave(
    file = paste(
      plot_dir,
      paste0("tasks_over_time_", trt, ".pdf"),
      sep = "/"
    ),
    plot = p,
    width = 20,
    height = 40
  )
}
```

## Symbiont interaction values

Load symbiont interaction values

```{r}
sym_int_vals_path <-  paste(
  working_directory,
  "data",
  "symbiont_interaction_values.csv",
  sep = "/"
)
sym_int_vals_data <- read_csv(sym_int_vals_path)

sym_int_vals_data <- sym_int_vals_data %>%
  mutate(
    SEED = as.factor(SEED),
    START_MOI = as.factor(START_MOI),
    VERTICAL_TRANSMISSION = as.factor(VERTICAL_TRANSMISSION),
    MUTUALIST_CYCLE_DONATE_MULTIPLIER = as.factor(MUTUALIST_CYCLE_DONATE_MULTIPLIER), # nolint: line_length_linter.
    PARASITE_CYCLE_STEAL_MULTIPLIER = as.factor(PARASITE_CYCLE_STEAL_MULTIPLIER), # nolint: line_length_linter.
    HEALTH_TYPE = as.factor(HEALTH_TYPE),
    PARASITE_CYCLE_LOSS_PROP = as.factor(PARASITE_CYCLE_LOSS_PROP),
    HEALTH_INTERACTION_CHANCE = as.factor(HEALTH_INTERACTION_CHANCE),
    PARASITE_BASE_CYCLE_PROP = as.factor(PARASITE_BASE_CYCLE_PROP)
  ) %>%
  mutate(
    treatment = paste0(
      "VT=", VERTICAL_TRANSMISSION
    ),
    interaction_multiplier = as.factor(MUTUALIST_CYCLE_DONATE_MULTIPLIER)
  ) %>%
  mutate(
    treatment = as.factor(
      treatment
    ),
    # Note: could come back to this to adjust how we do sorting
    super_mutualist_prop = (Hist_0.8 + Hist_0.9) / count
  )

# Give each seed a rep_id based on its super mutualist prop
sym_int_vals_data <- sym_int_vals_data %>%
  group_by(
    treatment
  ) %>%
  arrange(
    super_mutualist_prop
  ) %>%
  mutate(
    rep_id = rep(1:50, 1)
  ) %>%
  ungroup()

```

Pivot on histogram columns to long format

```{r}
long_sym_int_vals_data <- sym_int_vals_data %>%
  select(
    SEED,
    VERTICAL_TRANSMISSION,
    treatment,
    count,
    rep_id,
    super_mutualist_prop,
    "Hist_-0.1",
    "Hist_-0.2",
    "Hist_-0.3",
    "Hist_-0.4",
    "Hist_-0.5",
    "Hist_-0.6",
    "Hist_-0.7",
    "Hist_-0.8",
    "Hist_-0.9",
    "Hist_-1",
    "Hist_0.0",
    "Hist_0.1",
    "Hist_0.2",
    "Hist_0.3",
    "Hist_0.4",
    "Hist_0.5",
    "Hist_0.6",
    "Hist_0.7",
    "Hist_0.8",
    "Hist_0.9"
  ) %>%
  rename(total_syms = count) %>%
  pivot_longer(
    cols = -c(
      SEED,
      VERTICAL_TRANSMISSION,
      treatment,
      total_syms,
      super_mutualist_prop,
      rep_id
    ),
    names_to = "int_val_range",
    values_to = "int_val_count"
  ) %>%
  mutate(
    int_val_range = as.factor(int_val_range)
  )
```

Summarize interaction value ranges:
- Collapse bins to fewer categories, matching with ALife 2025
  tag paper bins.

```{r}
long_sym_int_vals_data <- long_sym_int_vals_data %>%
  # Recategorize
  mutate(
    int_val_range_reduced = case_when(
      int_val_range == "Hist_-1" ~ "-1.0 to -0.8 (Parasitic)",
      int_val_range == "Hist_-0.9" ~ "-1.0 to -0.8 (Parasitic)",
      int_val_range == "Hist_-0.8" ~ "-0.8 to -0.6 (Parasitic)",
      int_val_range == "Hist_-0.7" ~ "-0.8 to -0.6 (Parasitic)",
      int_val_range == "Hist_-0.6" ~ "-0.6 to -0.4 (Detrimental)",
      int_val_range == "Hist_-0.5" ~ "-0.6 to -0.4 (Detrimental)",
      int_val_range == "Hist_-0.4" ~ "-0.4 to -0.2 (Detrimental)",
      int_val_range == "Hist_-0.3" ~ "-0.4 to -0.2 (Detrimental)",
      int_val_range == "Hist_-0.2" ~ "-0.2 to 0 (Nearly neutral)",
      int_val_range == "Hist_-0.1" ~ "-0.2 to 0 (Nearly neutral)",
      int_val_range == "Hist_0.0" ~ "0 to 0.2 (Nearly neutral)",
      int_val_range == "Hist_0.1" ~ "0 to 0.2 (Nearly neutral)",
      int_val_range == "Hist_0.2" ~ "0.2 to 0.4 (Positive)",
      int_val_range == "Hist_0.3" ~ "0.2 to 0.4 (Positive)",
      int_val_range == "Hist_0.4" ~ "0.4 to 0.6 (Positive)",
      int_val_range == "Hist_0.5" ~ "0.4 to 0.6 (Positive)",
      int_val_range == "Hist_0.6" ~ "0.6 to 0.8 (Mutualistic)",
      int_val_range == "Hist_0.7" ~ "0.6 to 0.8 (Mutualistic)",
      int_val_range == "Hist_0.8" ~ "0.8 to 1.0 (Mutualistic)",
      int_val_range == "Hist_0.9" ~ "0.8 to 1.0 (Mutualistic)",
      .default = "unknown"
    )
  ) %>%
  mutate(
    int_val_range_reduced = factor(
      int_val_range_reduced,
      levels = c(
        "-1.0 to -0.8 (Parasitic)",
        "-0.8 to -0.6 (Parasitic)",
        "-0.6 to -0.4 (Detrimental)",
        "-0.4 to -0.2 (Detrimental)",
        "-0.2 to 0 (Nearly neutral)",
        "0 to 0.2 (Nearly neutral)",
        "0.2 to 0.4 (Positive)",
        "0.4 to 0.6 (Positive)",
        "0.6 to 0.8 (Mutualistic)",
        "0.8 to 1.0 (Mutualistic)"
      ),
      labels = c(
        "-1.0:-0.8 (Parasitic)",
        "-0.8:-0.6 (Parasitic)",
        "-0.6:-0.4 (Detrimental)",
        "-0.4:-0.2 (Detrimental)",
        "-0.2:0.0 (Nearly neutral)",
        "0.0:0.2 (Nearly neutral)",
        "0.2:0.4 (Positive)",
        "0.4:0.6 (Positive)",
        "0.6:0.8 (Mutualistic)",
        "0.8:1.0 (Mutualistic)"
      )
    )
  ) %>%
  # Collapse
  group_by(
    SEED,
    total_syms, # Note: total syms will be same for all rows of same seed.
    VERTICAL_TRANSMISSION,
    treatment,
    int_val_range_reduced,
    rep_id
  ) %>%
  summarize(
    int_val_count = sum(int_val_count)
  ) %>%
  mutate(
    int_val_prop = int_val_count / total_syms,
    VT = VERTICAL_TRANSMISSION
  )



```

```{r}
tenhelix <- c(
  "#891901",
  "#B50142",
  "#D506AD",
  "#AB08FF",
  "#5731FD",
  "#4755FF",
  "#42bcf5",
  "#42e0f5",
  "#9af1fc",
  "#c5f6fc"
)
sym_int_val_props_plt <- long_sym_int_vals_data %>%
  ggplot(
    mapping = aes(
      fill = int_val_range_reduced,
      x = rep_id,
      y = int_val_count,
      na.rm = TRUE
    )
  ) +
  geom_col(
    position = "fill",
    na.rm = TRUE,
    width = 1
  ) +
  scale_fill_manual(
    name = "Behavior",
    values = tenhelix
  ) +
  labs(
    x = "Replicate",
    y = "Proportion of symbiont population"
  ) +
  facet_wrap(
    ~ VT,
    nrow = 1,
    labeller = label_both
  ) +
  guides(
    fill = guide_legend(nrow = 2, byrow = TRUE)
  ) +
  theme(
    legend.position = "bottom"
  )

ggsave(
  file = paste(plot_dir, "sym_interaction_value_counts.pdf", sep = "/"),
  plot = sym_int_val_props_plt,
  width = 30,
  height = 20
)
```

```{r}
temp <- sym_int_vals_data %>%
  arrange(mean_intval)
```

## Phenotypic diversity

### Host diversity

#### Host number of task sets

```{r}
host_parent_num_task_sets_plt <- time_data %>%
  filter(update %% 500 == 0) %>%
  ggplot(
    aes(
      x = update,
      y = CurUpdate_host_parent_num_task_sets,
      color = VERTICAL_TRANSMISSION,
      fill = VERTICAL_TRANSMISSION
    )
  ) +
  stat_summary(fun = "mean", geom = "line") +
  stat_summary(
    fun.data = "mean_cl_boot",
    fun.args = list(conf.int = 0.95),
    geom = "ribbon",
    alpha = 0.2,
    linetype = 0
  ) +
  theme(
    legend.position = "bottom"
  )
ggsave(
  file = paste(plot_dir, "host_parent_num_task_sets_plt.pdf", sep = "/"),
  plot = host_parent_num_task_sets_plt,
  width = 15,
  height = 15
)
```

#### Host entropy of task sets

```{r}
host_parent_entropy_task_sets_plt <- time_data %>%
  filter(update %% 500 == 0) %>%
  ggplot(
    aes(
      x = update,
      y = CurUpdate_host_parent_entropy_task_sets,
      color = VERTICAL_TRANSMISSION,
      fill = VERTICAL_TRANSMISSION
    )
  ) +
  stat_summary(fun = "mean", geom = "line") +
  stat_summary(
    fun.data = "mean_cl_boot",
    fun.args = list(conf.int = 0.95),
    geom = "ribbon",
    alpha = 0.2,
    linetype = 0
  ) +
  theme(
    legend.position = "bottom"
  )
ggsave(
  file = paste(plot_dir, "host_parent_entropy_task_sets_plt.pdf", sep = "/"),
  plot = host_parent_entropy_task_sets_plt,
  width = 15,
  height = 15
)
```

### Symbiont diversity

#### Symbiont number of task sets

```{r}
symbiont_parent_num_task_sets_plt <- time_data %>%
  filter(update %% 500 == 0) %>%
  ggplot(
    aes(
      x = update,
      y = CurUpdate_sym_parent_num_task_sets,
      color = VERTICAL_TRANSMISSION,
      fill = VERTICAL_TRANSMISSION
    )
  ) +
  stat_summary(fun = "mean", geom = "line") +
  stat_summary(
    fun.data = "mean_cl_boot",
    fun.args = list(conf.int = 0.95),
    geom = "ribbon",
    alpha = 0.2,
    linetype = 0
  ) +
  theme(
    legend.position = "bottom"
  )
ggsave(
  file = paste(plot_dir, "symbiont_parent_num_task_sets_plt.pdf", sep = "/"),
  plot = symbiont_parent_num_task_sets_plt,
  width = 15,
  height = 15
)
```

#### Symbiont entropy of task sets

```{r}
symbiont_parent_entropy_task_sets_plt <- time_data %>%
  filter(update %% 500 == 0) %>%
  ggplot(
    aes(
      x = update,
      y = CurUpdate_sym_parent_entropy_task_sets,
      color = VERTICAL_TRANSMISSION,
      fill = VERTICAL_TRANSMISSION
    )
  ) +
  stat_summary(fun = "mean", geom = "line") +
  stat_summary(
    fun.data = "mean_cl_boot",
    fun.args = list(conf.int = 0.95),
    geom = "ribbon",
    alpha = 0.2,
    linetype = 0
  ) +
  theme(
    legend.position = "bottom"
  )
ggsave(
  file = paste(plot_dir, "symbiont_parent_entropy_task_sets_plt.pdf", sep = "/"),
  plot = symbiont_parent_entropy_task_sets_plt,
  width = 15,
  height = 15
)
```

## Generations elapsed

### Hosts

```{r}
host_avg_gens_elapsed_plt <- summary_data %>%
  ggplot(
    aes(
      x = VERTICAL_TRANSMISSION,
      fill = VERTICAL_TRANSMISSION,
      y = CurUpdate_host_mean_generations
    )
  ) +
  geom_point(
    mapping = aes(color = VERTICAL_TRANSMISSION),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_x_discrete() +
  theme(
    legend.position = "none",
    axis.text.x = element_text(
      angle = 60,
      hjust = 1
    )
  )
ggsave(
  file = paste(plot_dir, "host_avg_gens_elapsed_plt.pdf", sep = "/"),
  plot = host_avg_gens_elapsed_plt,
  width = 10,
  height = 10
)
```

### Symbionts

```{r}
sym_avg_gens_elapsed_plt <- summary_data %>%
  ggplot(
    aes(
      x = VERTICAL_TRANSMISSION,
      fill = VERTICAL_TRANSMISSION,
      y = CurUpdate_sym_mean_generations
    )
  ) +
  geom_point(
    mapping=aes(color = VERTICAL_TRANSMISSION),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_x_discrete() +
  theme(
    legend.position = "none",
    axis.text.x = element_text(
      angle = 60,
      hjust = 1
    )
  )
ggsave(
  file = paste(plot_dir, "sym_avg_gens_elapsed_plt.pdf", sep = "/"),
  plot = sym_avg_gens_elapsed_plt,
  width = 10,
  height = 10
)
```

## Task complexity as a function of mutualism proportion

```{r}
library(janitor)
# Tack on proportion of mutualists
cleaned_host_max_task_complexity_data <- clean_names(
  host_max_task_complexity_data
)
sym_int_vals_vs_complexity_data <- sym_int_vals_data %>%
  clean_names(replace = c("-" = "m")) %>%
  mutate(
    mutualist_prop = (hist_0_6 + hist_0_7 + hist_0_8 + hist_0_9) / count,
    parasite_prop = (hist_m1 + hist_m0_9 + hist_m0_8 + hist_m0_7) / count
  ) %>%
  left_join(
    cleaned_host_max_task_complexity_data,
    by = c("seed", "vertical_transmission")
  )
```

```{r}
sym_int_vals_vs_host_complexity_plt <- sym_int_vals_vs_complexity_data %>%
  filter(!is.na(mutualist_prop)) %>%
  ggplot(
    aes(
      x = mutualist_prop,
      y = max_task_complexity,
      color = vertical_transmission,
      fill = vertical_transmission
    )
  ) +
  geom_point()

sym_int_vals_vs_host_complexity_plt

stat_data <- sym_int_vals_vs_complexity_data %>%
  filter(!is.na(mutualist_prop))
cor.test(
  sym_int_vals_vs_complexity_data$mutualist_prop,
  sym_int_vals_vs_complexity_data$max_task_complexity,
  # formula = max_task_complexity ~ mutualist_prop,
  # data = stat_data,
  method = "spearman",
  exact = FALSE
)
```

```{r}
para_vs_host_complexity_plt <- sym_int_vals_vs_complexity_data %>%
  filter(!is.na(parasite_prop)) %>%
  ggplot(
    aes(
      x = parasite_prop,
      y = max_task_complexity,
      color = vertical_transmission,
      fill = vertical_transmission
    )
  ) +
  geom_point()

para_vs_host_complexity_plt

stat_data <- sym_int_vals_vs_complexity_data %>%
  filter(!is.na(parasite_prop))
cor.test(
  sym_int_vals_vs_complexity_data$parasite_prop,
  sym_int_vals_vs_complexity_data$max_task_complexity,
  # formula = max_task_complexity ~ mutualist_prop,
  # data = stat_data,
  method = "spearman",
  exact = FALSE
)
```




## Manuscript figure


```{r}
tenhelix <- c(
  "#891901",
  "#B50142",
  "#D506AD",
  "#AB08FF",
  "#5731FD",
  "#4755FF",
  "#42bcf5",
  "#42e0f5",
  "#9af1fc",
  "#c5f6fc"
)

sym_int_val_props_fig <- long_sym_int_vals_data %>%
  ggplot(
    mapping = aes(
      fill = int_val_range_reduced,
      x = rep_id,
      y = int_val_count,
      na.rm = TRUE
    )
  ) +
  geom_col(
    position = "fill",
    na.rm = TRUE,
    width = 1
  ) +
  scale_fill_manual(
    name = "Behavior",
    values = tenhelix
  ) +
  labs(
    x = "Replicate",
    y = "Symbiont population\nproportion"
  ) +
  facet_nested_wrap(~"Vertical transmission rate" + VT, nrow = 1) +
  guides(
    fill = guide_legend(nrow = 2, byrow = TRUE)
  ) +
  theme(
    legend.position = "bottom",
    legend.text = element_text(size = 10),
    legend.title = element_text(size = 10, angle = 90),
    strip.background = element_rect(fill = "white", colour = "black"),
    strip.text = element_text(size = 10),
    axis.text.x = element_blank(),
    axis.title.x = element_text(size = 10),
    axis.ticks.x = element_blank(),
    axis.text.y = element_text(size = 10),
    axis.title.y = element_text(size = 10)
  ) + panel_border(color = "black", size = 0.5)

ggsave(
  file = paste(plot_dir, "sym_int_val_props_fig.pdf", sep = "/"),
  plot = sym_int_val_props_fig,
  width = 10,
  height = 4
)
```